### 2012-05-11 Docs/Proposal/proposal.tex

\documentclass[letterpaper]{article}
\usepackage[margin=1.2in]{geometry}
\usepackage{ifthen}
\usepackage{stmaryrd}
\usepackage{mathpartir}
\usepackage[usenames]{color}
\usepackage{hyperref}
\usepackage{multicol}
\usepackage{fancyvrb}

% For HsColour
%\newsavebox{\spaceb}
%\newsavebox{\tabb}
%\savebox{\spaceb}[1ex]{~}
%\savebox{\tabb}[4ex]{~}
%\newcommand{\hsspace}{\usebox{\spaceb}}
%\newcommand{\hstab}{\usebox{\tabb}}
%\newcommand{\conceal}[1]{}

\input{pygments.tex}
\input{amsden-commands.tex}

\title{Push-Based Signal-Function Functional Reactive Programming (Thesis Proposal)}
\author{Edward Amsden}
\date{\today}

\begin{document}

\maketitle
\input{abstract.tex}
\input{introduction.tex}
\input{background.tex}
\input{proposed_work.tex}

\bibliographystyle{acm}
\bibliography{proposal}
\end{document}