### 2012-05-11 Docs/Proposal/sf-frp-example-hs.tex

\begin{Verbatim}[commandchars=\\\{\}]
\PY{c+c1}{-- Assumed primitives:}
\PY{c+c1}{-- arr :: (a -\PYZgt{} b) -\PYZgt{} SF a b}
\PY{c+c1}{-- \PYZgt{}\PYZgt{}\PYZgt{} :: SF a b -\PYZgt{} SF b c -\PYZgt{} SF a c}
\PY{c+c1}{-- first :: SF a b -\PYZgt{} SF (a, c) (b, c)}
\PY{c+c1}{-- second :: SF a b -\PYZgt{} SF (c, a) (c, b)}
\PY{c+c1}{-- switch :: SF a (b, Event (SF a b)) -\PYZgt{} SF a b}
\PY{c+c1}{-- reactimateWithMouseAndSound :: SF (Event (), ((Int, Int), Double)) (Double) -\PYZgt{} IO ()}

\PY{c+c1}{-- Assumed given signal functions:}
\PY{c+c1}{-- bandpass :: SF (Double, (Double, Double)) Double}

\PY{k+kr}{import} \PY{n+nn}{Control.AFRP} \PY{c+c1}{-- Our theoretical AFRP}

\PY{n+nf}{mouseCoordsToDoubles} \PY{o+ow}{::} \PY{k+kt}{SF} \PY{p}{(}\PY{k+kt}{Int}\PY{p}{,} \PY{k+kt}{Int}\PY{p}{)} \PY{p}{(}\PY{k+kt}{Double}\PY{p}{,} \PY{k+kt}{Double}\PY{p}{)}
\PY{n+nf}{mouseCoordsToDoubles} \PY{o+ow}{=} 
  \PY{n}{arr} \PY{p}{(}\PY{n+nf}{\PYZbs{}}\PY{p}{(}\PY{n}{x}\PY{p}{,} \PY{n}{y}\PY{p}{)} \PY{o+ow}{-\PYZgt{}} \PY{p}{(}\PY{n}{fromIntegral} \PY{n}{x} \PY{o}{/} \PY{n}{fromIntegral} \PY{n}{screenWidth}\PY{p}{,} 
                   \PY{n}{fromIntegral} \PY{n}{y} \PY{o}{/} \PY{n}{fromIntegral} \PY{n}{screenHeight}\PY{p}{)}\PY{p}{)}

\PY{n+nf}{highFreq} \PY{o+ow}{::} \PY{k+kt}{SF} \PY{k+kt}{Double} \PY{k+kt}{Double}
\PY{n+nf}{highFreq} \PY{o+ow}{=} \PY{n}{arr} \PY{p}{(}\PY{n+nf}{\PYZbs{}}\PY{n}{x} \PY{o+ow}{-\PYZgt{}} \PY{l+m+mi}{60} \PY{o}{*} \PY{n}{exp}\PY{p}{(} \PY{n}{x} \PY{o}{*} \PY{n}{ln} \PY{p}{(}\PY{l+m+mi}{20000} \PY{o}{/} \PY{l+m+mi}{60}\PY{p}{)}\PY{p}{)}\PY{p}{)}

\PY{n+nf}{lowFreq} \PY{o+ow}{::} \PY{k+kt}{SF} \PY{p}{(}\PY{k+kt}{Double}\PY{p}{,} \PY{k+kt}{Double}\PY{p}{)} \PY{k+kt}{Double}
\PY{n+nf}{lowFreq} \PY{o+ow}{=} \PY{n}{arr} \PY{p}{(}\PY{n+nf}{\PYZbs{}}\PY{p}{(}\PY{n}{high}\PY{p}{,} \PY{n}{y}\PY{p}{)} \PY{o+ow}{-\PYZgt{}} \PY{l+m+mi}{60} \PY{o}{*} \PY{n}{exp} \PY{p}{(}\PY{n}{y} \PY{o}{*} \PY{n}{ln} \PY{p}{(}\PY{n}{high} \PY{o}{/} \PY{l+m+mi}{60}\PY{p}{)}\PY{p}{)}\PY{p}{)}

\PY{n+nf}{lowHighFreq} \PY{o+ow}{::} \PY{k+kt}{SF} \PY{p}{(}\PY{k+kt}{Double}\PY{p}{,} \PY{k+kt}{Double}\PY{p}{)} \PY{p}{(}\PY{k+kt}{Double}\PY{p}{,} \PY{k+kt}{Double}\PY{p}{)}
\PY{n+nf}{lowHighFreq} \PY{o+ow}{=} \PY{n}{first} \PY{n}{highFreq} \PY{o}{\PYZgt{}\PYZgt{}\PYZgt{}} 
              \PY{n}{arr} \PY{p}{(}\PY{n+nf}{\PYZbs{}}\PY{p}{(}\PY{n}{high}\PY{p}{,} \PY{n}{y}\PY{p}{)} \PY{o+ow}{-\PYZgt{}} \PY{p}{(}\PY{n}{high}\PY{p}{,} \PY{p}{(}\PY{n}{high}\PY{p}{,} \PY{n}{y}\PY{p}{)}\PY{p}{)}\PY{p}{)} \PY{o}{\PYZgt{}\PYZgt{}\PYZgt{}}
              \PY{n}{second} \PY{n}{lowFreq}

\PY{n+nf}{mouseCoordsToLowHighFreqs} \PY{o+ow}{::} \PY{k+kt}{SF} \PY{p}{(}\PY{k+kt}{Int}\PY{p}{,} \PY{k+kt}{Int}\PY{p}{)} \PY{p}{(}\PY{k+kt}{Double}\PY{p}{,} \PY{k+kt}{Double}\PY{p}{)} 
\PY{n+nf}{mouseCoordsToLowHighFreqs} \PY{o+ow}{=} \PY{n}{mouseCoordsToDoubles} \PY{o}{\PYZgt{}\PYZgt{}\PYZgt{}} \PY{n}{lowHighFreq}

\PY{n+nf}{toggleFilter} \PY{o+ow}{=} 
  \PY{k+kr}{let} \PY{n}{filterOn} \PY{o+ow}{=} 
        \PY{n}{switch} \PY{p}{(}\PY{n}{arr} \PY{p}{(}\PY{n+nf}{\PYZbs{}}\PY{p}{(}\PY{n}{evt}\PY{p}{,} \PY{p}{(}\PY{n}{mouse}\PY{p}{,} \PY{n}{audio}\PY{p}{)}\PY{p}{)} \PY{o+ow}{-\PYZgt{}} 
                        \PY{p}{(}\PY{p}{(}\PY{n}{audio}\PY{p}{,} \PY{n}{mouse}\PY{p}{)}\PY{p}{,} \PY{n}{fmap} \PY{p}{(}\PY{n}{const} \PY{n}{filterOff}\PY{p}{)} \PY{n}{evt}\PY{p}{)}\PY{p}{)} \PY{o}{\PYZgt{}\PYZgt{}\PYZgt{}}
                \PY{n}{first} \PY{p}{(}\PY{n}{second} \PY{n}{mouseCoordsToLowHighFreqs} \PY{o}{\PYZgt{}\PYZgt{}\PYZgt{}} \PY{n}{bandpass}\PY{p}{)}\PY{p}{)}
      \PY{n}{filterOff} \PY{o+ow}{=} \PY{n}{switch} \PY{p}{(}\PY{n}{arr} \PY{p}{(}\PY{n+nf}{\PYZbs{}}\PY{p}{(}\PY{n}{evt}\PY{p}{,} \PY{p}{(}\PY{k+kr}{\PYZus{}}\PY{p}{,} \PY{n}{audio}\PY{p}{)}\PY{p}{)} \PY{o+ow}{-\PYZgt{}}
                                \PY{p}{(}\PY{n}{audio}\PY{p}{,} \PY{n}{fmap} \PY{p}{(}\PY{n}{const} \PY{n}{filterOn}\PY{p}{)} \PY{n}{evt}\PY{p}{)}\PY{p}{)}\PY{p}{)}
  \PY{k+kr}{in} \PY{n}{filterOn}

\PY{n+nf}{main} \PY{o+ow}{::} \PY{k+kt}{IO} \PY{n+nb}{()}
\PY{n+nf}{main} \PY{o+ow}{=} \PY{n}{reactimateWithMouseAndSound} \PY{n}{toggleFilter}
\end{Verbatim}