### 2012-05-11 Docs/Proposal/signalvectors-hs.tex

\begin{Verbatim}[commandchars=\\\{\}]
\PY{c+c1}{-- Signal Vectors}
\PY{k+kr}{data} \PY{k+kt}{SVEmpty}
\PY{k+kr}{data} \PY{k+kt}{SVSignal} \PY{o+ow}{::} \PY{o}{*} \PY{o+ow}{-\PYZgt{}} \PY{o}{*}
\PY{k+kr}{data} \PY{k+kt}{SVEvent}  \PY{o+ow}{::} \PY{o}{*} \PY{o+ow}{-\PYZgt{}} \PY{o}{*}
\PY{k+kr}{data} \PY{k+kt}{SVAppend} \PY{o+ow}{::} \PY{o}{*} \PY{o+ow}{-\PYZgt{}} \PY{o}{*} \PY{o+ow}{-\PYZgt{}} \PY{o}{*}

\PY{c+c1}{-- Signal Vector indexes}
\PY{k+kr}{data} \PY{k+kt}{SVIndex} \PY{o+ow}{::} \PY{p}{(}\PY{o}{*} \PY{o+ow}{-\PYZgt{}} \PY{o}{*}\PY{p}{)} \PY{o+ow}{-\PYZgt{}} \PY{o}{*} \PY{o+ow}{-\PYZgt{}} \PY{o}{*} \PY{k+kr}{where}
  \PY{k+kt}{SVISignal} \PY{o+ow}{::} \PY{n}{a} \PY{o+ow}{-\PYZgt{}} \PY{k+kt}{SVIndex} \PY{n}{p} \PY{p}{(}\PY{k+kt}{SVSignal} \PY{n}{p} \PY{n}{a}\PY{p}{)}
  \PY{k+kt}{SVIEvent}  \PY{o+ow}{::} \PY{n}{a} \PY{o+ow}{-\PYZgt{}} \PY{k+kt}{SVIndex} \PY{n}{p} \PY{p}{(}\PY{k+kt}{SVEvent} \PY{n}{p} \PY{n}{a}\PY{p}{)}
  \PY{k+kt}{SVILeft}   \PY{o+ow}{::} \PY{k+kt}{SVIndex} \PY{n}{p} \PY{n}{svl} \PY{o+ow}{-\PYZgt{}} \PY{k+kt}{SVIndex} \PY{n}{p} \PY{p}{(}\PY{k+kt}{SVAppend} \PY{n}{svl} \PY{n}{svr}\PY{p}{)}
  \PY{k+kt}{SVIRight}  \PY{o+ow}{::} \PY{k+kt}{SVIndex} \PY{n}{p} \PY{n}{svr} \PY{o+ow}{-\PYZgt{}} \PY{k+kt}{SVIndex} \PY{n}{p} \PY{p}{(}\PY{k+kt}{SVAppend} \PY{n}{svl} \PY{n}{svr}\PY{p}{)}
\end{Verbatim}