### 2012-05-12 Docs/HaskellSymposium2012/frp.tex

\section{Signal-Function FRP}
\label{section:functional_reactive_programming}

Most reactive programs are written in an imperative style, using low-level and non-composable abstractions including callbacks
or object based event handlers, called by event loops. This ties the model of interactivity to low-level implementation details such as timing and event handling models.

FRP implies that a model should keep the characteristics of functional programming (i.e. that the basic constructs of the language
should remain first-class) while incorporating reactivity into the language model. In particular, functions should be lifted to operate on reactive values,
and functions themselves ought to be reactive.

The goal of FRP is to provide compositional and high-level abstractions for creating reactive programs. The key
abstractions of FRP are behaviors or signals, which are time-dependent values defined at every point in continuous time, and events, which are 
values defined at countably many points in time. An FRP system will provide functions to manipulate events and signals and to react
to events by replacing a portion of the running program in response to an event. Behaviors and events, or some abstraction
based on them, will be first class, in keeping with the spirit of functional programming. Programs implemented in functional reactive
models should of course be efficiently executable. This has proven to be the main challenge in implementing Functional Reactive Programming.

The two general approaches to FRP are ``classic'' FRP, where behaviors and signals are first-class and reactive objects, and ``signal-function'' FRP,
where transformers of signals and events are first-class and reactive objects.

In signal-function FRP, programs are written by composing signal functions, which take one or more signals and/or event streams as input
and produce one or more signals and/or event-streams as output. Ideally, the programmer has only point-wise access to signals (values at the current
point in time) and conditional access to event streams (access when there is an event occurrence).

In current signal-function FRP systems, such as Yampa~\cite{Nilsson2005}, a signal function transforms a single signal,
and combinators combine signal functions by splitting input and/or combining output values. Event streams
are represented as option-valued components of the signal. This leads to inefficient implementation, as event streams
are sampled at every time step. Further, the latency between event occurrent and event response is, on average, one-half
of the sampling rate~\cite{Elliott2009}. The pull-based evaluation is necessitated by the single signal value, since
there is, in general, no way to separate components of a signal which have been updated or do carry event occurrences
from those which have not.

In N-Ary FRP~\cite{Sculthorpe2011}, signal functions conceptually operate on {\em signal vectors}, which represent collections
of signals and event streams at the type level. This opens the possibility of push-based evaluation, since there
is now a clear distinction between individual signals and event streams. A push-based evaluation of an N-Ary FRP
system can define reactive operations for signal functions in terms of individual inputs, with combinators
maintaining the distinctions between elements of the signal vector and skipping evaluation of signal functions for
which there is no input.