### 2012-05-12 Docs/HaskellSymposium2012/implementation.tex

\section{Implementation}
\label{section:implementation}

Since the signal functions will operate on signal vectors, the first step in an implementation is to represent signal vectors.
We represent signal vectors at the type level as empty types:

\begin{verbatim}
-- An empty signal vector
data SVEmpty
-- A singleton vector of a signal
data SVSignal a
-- A singleton vector of an event stream
data SVEvent a
-- The concatenation of two signal vectors
data SVAppend svl svr 
\end{verbatim}

Once we have this type-level representation of signal vectors, we represent components of the signal vector as a GADT\footnote{Generalized Algebraic Datatype}
with phantom types:

\begin{verbatim}
data SVIndex ::   (* -> *) -> * -> * where
  SVISignal  ::   tc a -> SVIndex tc (SVSignal a)
  SVIEvent   ::   tc a -> SVIndex tc (SVEvent a)
  SVILeft    ::   SVIndex tc svl 
               -> SVIndex tc (SVAppend svl svr)
  SVIRight   ::   SVIndex tc svr 
               -> SVIndex tc (SVAppend svl svr)
\end{verbatim}

Note the addition of the higher-kinded type parameter, which allows us to represent components where the types of the data
in the component are dependent on but not equivalent to the the types of the signals and events in the vector. For instance,
we may instantiate $tc$ as $(\rightarrow IO ())$ to represent a new handler for the output of a signal function.

Another datatype provides a means of accumulating partial information about the state of a signal vector:
\begin{verbatim}
data SVMemory ::   (* -> *) -> * -> *
  SVMSignal   ::   tc a -> SVMemory tc (SVSignal a)
  SVMEvent    ::   tc a -> SVMemory tc (SVEvent a)
  SVMNothing  ::   SVMemory tc sv
  SVMBoth     ::   SVMemory tc svl
                -> SVMemory tc svr 
                -> SVMemory tc (SVAppend svl svr)
\end{verbatim}

Again notice the higher-kinded type parameter, which in this case might be used to produce a record of the current
handlers for the outputs of a signal function.

With these representations in hand we can describe an implementation type for signal functions:
\begin{verbatim}
newtype Id a = Id a

data NonInitialized
data Initialized

data SF  :: * -> * -> * -> * where
  SFInit ::  (SVIndex Id svIn 
              -> ([SVIndex Id svOut],
                  SF Initialized svIn svOut)) 
           -> SF Initialized svIn svOut
  SF     ::  (SVMemory svIn 
              -> (SVMemory Id svOut,
                  SF Initialized svInt svOut)) 
           -> SF NonInitialized svIn svOut
\end{verbatim}

Combinators produce the second constructor, which takes a partial memory of the input state of the signal function.
This enables preservation of the values of signals when a new signal function becomes a part of the network. 
Once a signal function has been supplied with the current state, it exists as a continuation (wrapped by the first constructor)
which takes a change (represented as an event occurrences or a new values for a signal in a signal vector index) to 
a list of output changes and a new signal function.

This implementation strategy is very similar to that used for other signal-function FRP systems, such as Yampa~\cite{Nilsson2005}.
The distinction is that, with the inputs separated, the continuation produced by a combinator may select the appropriate continuation
for a particular input, and components of the input may be responded to as they arrive, rather than all together at each time step.

The implementation strategy can be demonstrated with a few examples of combinators from the library:

The {\tt identity} signal function is perhaps the simplest. It is completely polymorphic in its input, and simply
copies inputs to its output:
\begin{verbatim}
identity :: SF NonInitialized sv sv
identity = SF (\mem -> (mem, identityInit))

identityInit :: SF Initialized sv sv
identityInit = 
  SFInit (\change -> ([change], identityInit))
\end{verbatim}

Note that every signal function must specify its replacement along with its output.
In this case, since identity does not react to but merely passes along inputs, its replacement is itself.

The sequential composition operator ({\tt >>>}) is perhaps the most crucial combinator.

The parallel composition functions {\tt first} and {\tt second} allow signal functions to
operate on one side of a signal vector, with the other side passed through unchanged.
{\tt first} is shown here.

The feedback combinator {\tt loop} allows a signal function to take its own input as output.

The remaining combinators are implemented similarly.