### 2012-06-20 Docs/Thesis/abstract.tex
Initial thesis layout

Functional Reactive Programming (FRP) is a promising family of abstractions for
encoding reactive and time-dependent programs in functional languages.
Signal-function FRP is a subfamily of FRP which provides {\em signal functions}
as first-class abstractions to the programmer, rather than signals and events as
is common in ``classic'' FRP systems.