### 2012-06-20 Docs/Thesis/background.tex

\chapter{Background}
\label{chapter:Background}

The concept of FRP is traced back to the Fran system for declarative 
animation~\cite{Elliott1997}.