### 2012-06-20 Docs/Thesis/comparison.tex

\chapter{Evaluation and Comparisons}
\label{chapter:Evaluation_and_Comparisons}

We have described the implementation of a push-based signal-function FRP system.
A push-based system should theoretically have advantages in event latency and
overall performance over a pull-based system. Further, we would like to compare
the evaluation interface in the current widely-used system with the one proposed
here, again in order to verify that the theoretical benefits do appear in
practice.