### 2012-06-20 Docs/Thesis/conclusions-and-further-work.tex

\chapter{Conclusions and Further Work}
\label{chapter:Conclusions_and_Further_Work}
I have presented TimeFlies, a push-based signal-function FRP system. I have
demonstrated that TimeFlies does realize the theoretical benefits of a
push-based signal-function system.