### 2012-06-20 Docs/Thesis/implementation.tex

\chapter{Implementation}
\label{chapter:Implemenation}

Having established our method of modeling reactivity, we proceed to a purely 
functional implementation of this method. For this implementation, we will make
significant use of certain advanced extensions to the Haskell type system.