### 2012-06-20 Docs/Thesis/introduction.tex

\chapter{Introduction}
\label{chapter:Introduction}

It is common when considering computation to simplify programs to simple
functions from one input to one output. Alas, this simplication generally does
not hold for even the simplest applications of programming. Most applications
demand that a program respond to multiple inputs and produce multiple outputs
scattered through time.

The common model for such programs consists of an event loop which maintains a
queue of input events. These events are then fed to event handlers which perform
various observable actions and mutate the state of the application.