### 2012-06-23 Docs/Thesis/implementation-signalfunctions.tex

\section{Signal Functions}
\label{section:Implementation-Signal_Functions}

To encode signal functions, we must create a compositional way to express
both the output and replacement responses to inputs. We must also have a means
of specifying the initial input and output signal of the signal function.

The interface types for signal functions are given in 
Figure~\ref{figure:signal-function-interface-types}.

\begin{figure}
\begin{minipage}[b]{0.5\linewidth}
\begin{code}
-- Signal memories
data SMemory p sv

-- Signal vector indices
data SVIndex p sv

-- Initialization types
data Initialized
data NonInitialized
\end{code}
\end{minipage}
\hspace{0.5cm}
\begin{minipage}[b]{0.5\linewidth}
\begin{code}
-- Signal vectors
data SVEmpty
data SVSignal a
data SVEvent a
data SVAppend svLeft svRight

-- Signal functions
data SF init svIn svOut
\end{code}
\end{minipage}
\vspace{\parskip}
\hrule
\caption{Interface types for signal functions}
\label{figure:signal-function-interface-types}
\end{figure}

Our implementation makes use of two value representations of signal vectors.
A signal memory is typed using a signal vector and carries values for a
(possibly empty) subset of the points in the signal vector. A signal vector
index is typed using a signal vector and carries a value for exactly one point
in the signal vector. In order to define memories and indices over signal
vectors, we make use of a type extension known as
Generalized~Algebraic~Datatypes~(GADTs)~\cite{Cheney2003,Xi2003}, which is
available in recent versions of the
Glasgow~Haskell~Compiler~(GHC)~\cite{PeytonJones2006}. The key distinction
of GADTs from standard Algebraic~Datatypes is that GADTs enable 
{\em type refinement}. A type constructor in a GADT specifies the specific type
with which each type parameter of the GADT's type constructor is instantiated.
This type is defined in terms of the types of the constructor's data members.
Any type variable which occurs in the type of a data member but does not occur
in the type of the constructor is existentially quantified\footnote{Existential
quantification, in the context of Algebraic Datatypes, is a means of ``hiding''
type variables in data member types. A function which pattern matches on a
constructor with existential quantification may not assume any instantiation of
the existentially quantified variable, but only typeclass constraints specified
in the constructor definition, and equality between identical type variables in
data members of the same constructor.}. GADTs generalize Algebraic Datatypes 
(ADTs) because an ADT may be translated to a GADT by simply using each unique
type variable in the types of the data members of a constructor to instantiate
the type parameter. As with ADTs, a type variable which appears in the type of a
data constructor but not in any data members may be arbitrarily instantiated.

The ability to specialize types is useful both for creating specialized
constructors for unique cases of the type parameters (e.g. for
optimization~\cite{Nilsson2005}) and for encoding properties at the type level
using empty types. We employ the second use here.% and the first later in Section~\ref{subsection:Implementation-Signal_Functions-Representation} when we describe the representation of signal functions.

Both signal memories and signal vector indices are GADTs:

\begin{code}
-- Signal memories
data SMemory :: (* -> *) -> * -> * where
  SMEmpty  :: SMemory p sv
  SMSignal :: p a -> SMemory p (SVSignal a)
  SMEvent  :: p a -> SMemory p (SVEvent a)
  SMBoth   :: SMemory p svLeft -> SMemory p svRight 
           -> SMemory p (SVAppend svLeft svRight)

-- Signal vector indices
data SVIndex :: (* -> *) -> * -> * where
  SVISignal :: p a -> SVIndex p (SVSignal a)
  SVIEvent  :: p a -> SVIndex p (SVEvent a)
  SVILeft   :: SVIndex p svLeft -> SVIndex p (SVAppend svLeft svRight)
  SVIRight  :: SVIndex p svRight -> SVIndex p (SVAppend svLeft svRight)

-- ``Identity'' newtype
newtype Id a = Id a
\end{code}

The higher-kind type parameter {\tt p} in both type constructors is a
convenience. It allows us to provide a type constructor which will be mapped
over the types of signals and events in the signal vector to produce the actual
type of values stored. This ability is used in
Section~\ref{section:Implementation-Evaluation_Interface} (Evaluation Interface),
permitting the memory type, in particular, to carry functions from the types of
points in the signal vector to the monadic action used to handle the
corresponding outputs.

Next we consider the type for signal functions. A signal function must be able
to have an initial signal input specified, and must be able to produce output
signals, output events, and replace itself, based on time steps and input
signals or event occurrences.

The definition of the {\tt SF} datatype is:

\begin{code}
data SF :: * -> * -> * -> * where
  SF     :: (SMemory Id svIn -> (SMemory Id svOut, SF Initialized svIn svOut))
         -> SF NonInitialized svIn svOut
  SFInit :: (Double -> SMemory Id svIn 
             -> (SMemory Id svOut, 
                 [SVIndex Id svOut],
                 SF Initialized svIn svOut))
         -> (SVIndex Id svIn 
             -> ([SVIndex Id svOu], SF Initialized svIn svOut))
         -> SF Initialized svIn svOut
\end{code}

The first constructor is used in the user-facing combinators to construct
non-initialized signal functions. It contains a single function from an intial
signal (which may contain empty values) to an initial output signal and
an initialized signal function with the same input and output types.

The second constructor is used in the implementation of both the non-initialized
and initialized signal functions to construct initialized signal functions. It
takes two functions. The first is a function from a time step and an updated
signal to an updated output signal, a set of output event occurrences, and a new
signal function. The second is a function from an event occurrence to a set
of event occurrences and a new signal function.

\subsection{Combinators}
\label{subsection:Implementation-Signal_Functions-Combinators}
From the perspective of the library user, signal functions are built from a set
of combinators which produce or modify signal functions. These combinators
include a set of primitive signal functions, a set of lifting functions to
produce signal functions from functions, a switching combinator,
time and integration combinators, and a set of routing combinators. A list
of the exported combinators is given in
Figure~\ref{figure:signal_function_combinators}.

\begin{figure}
\begin{code}
-- Simple signal functions
identity :: SignalFunction NonInitialized sv sv
constant :: a -> SignalFunction NonInitialized SVEmpty (SVSignal a)

-- Composition and routing
(>>>) :: SignalFunction NonInitialized svIn svBetween
      -> SignalFunction NonInitialized svBetween svOut
      -> SignalFunction NonInitialized svIn svOut

first :: SignalFunction NonInitialized svIn svOut
      -> SignalFunction NonInitialized (SVAppend svIn sv) (SVAppend svOut sv)

second :: SignalFunction NonInitialized svIn svOut
       -> SignalFunction NonInitialized (SVAppend sv svIn) (SVAppend sv svOut)

swap :: SignalFunction NonInitialized 
          (SVAppend svLeft svRight) (SVAppend svRight svLeft)
copy :: SignalFunction NonInitialized sv (SVAppend sv sv)
drop :: SignalFunction NonInitialized sv SVEmpty
\end{code}
\hrule
\caption{The combinators exported by the signal function interface}
\label{figure:signal_function_combinators}
\end{figure}

The simplest combinator is the {\tt identity} signal function. As its name would
suggest, the identity signal function simply replicates its input on its output.
It is implemented as:

\begin{code}
identity :: SF NonInitialized sv sv
identity = SF (\mem -> (identityInit, mem))

identityInit :: SF Initialized sv sv
identityInit = let sf = SFInit (\_ mem -> (mem, [], sf))
                               (\idx -> ([idx], sf))
               in sf
\end{code}

Only the first function is exported. In general, we use the convention that
the initialized version of a signal function is named with the signal function
name with {\tt Init} appended, and is not exported as part of the interface.