### 2012-06-23 Docs/Thesis/implementation.tex

\chapter{Implementation}
\label{chapter:Implementation}

Having established the semantics of the reactive system, I present a purely 
functional implementation. For this implementation, I will make significant use
of extensions to the Haskell type system. Properly used, these extensions allow
greater expressive flexibility to define our types by permitting us to encode
additional properties of values in their types.

\input{implementation-signalfunctions}

\input{implementation-evalinterface}