### 2012-06-28 Docs/Thesis/system_design_and_interface.tex
Change chapter semantics->system design and interface

\chapter{System Design and Interface}
\label{chapter:System_Design_and_Interface}

We have explored the current design space of FRP systems, and identified
strengths and weaknesses with the existing state-of-the-art systems. By
combining the best characteristics of both state-of-the-art systems, we hope
to produce an FRP system which combines the efficiency of push-based FRP system
with the modularity and causality safety of a signal-function FRP system.

The advantage of a push-based system is that it evaluates only those parts
of the network which are necessary to react to the input. In order to achieve
this behavior with signal functions, we must have a means of allowing the system
to distinguish between multiple inputs to a signal function. This is where the
concept of signal vectors from N-Ary FRP will be useful.