### 2012-10-03 Docs/IFL2012/presentation/presentation.tex

\documentclass{beamer}
\usetheme{JuanLesPins}
\usecolortheme{seahorse}

\usepackage{fancyvrb}
\usepackage{graphicx}
\usepackage[style=authortitle-icomp]{biblatex}

%gets rid of bottom navigation bars
\setbeamertemplate{footline}[]{}

%gets rid of navigation symbols
\setbeamertemplate{navigation symbols}{}

\bibliography{thesis}

\title{Push-Pull Signal-Function Functional Reactive Programming}
\author{Edward Amsden}
\institute{Rochester Institute of Technology}
\date{IFL 2012\\ \today}

\begin{document}

\AtBeginSection[]
{
\begin{frame}
\end{frame}
}

\begin{frame}[plain]
\titlepage
\includegraphics[width=9em]{RIT_GCCIS1}
\end{frame}

\section{Introduction}

\begin{frame}
\frametitle{Functional Reactive Programming (FRP)}
    \begin{itemize}
        \item {\em Functional}: First-class and higher-order functions.
        \item {\em Reactive}: Behavior changes in response to temporal inputs.
        \item Basic abstractions:
        \begin{itemize}
            \item Signals - Functions of time\footnote{Often also called behaviors.}.
            \item Events - Sequences of temporally ordered and labeled\\
                  discrete values.
        \end{itemize}
    \item ``Classic'' FRP - Signals and events are first-class.
    \end{itemize}
\end{frame}

\begin{frame}
\frametitle{Signal-Function FRP (SF-FRP)}
    \begin{itemize}
        \item Signal functions are transformers of events and signals.
        \item Signal functions are first-class in Signal-Function FRP.
        \item Signals and events are not first-class in Signal-Function FRP.
        \item This approach is more composable than first-class\\
              signals and events (since input may be transformed)\footcite{Courtney2001-1}.
    \end{itemize}
\end{frame}

\begin{frame}
\frametitle{Push vs. Pull Evaluation}
    \begin{itemize}
        \item When to evaluate what?
        \item Pull evaluation (``demand-driven''):\\ Evaluate when output is needed.
        \item Push evaluation (``data-driven''):\\ Evaluate when input is available.
        \item Ideally, an FRP system uses push for events and\\
              pull for signals\footcite{Elliott2009}.
        \item This has been achieved for ``classic'' FRP
              (first class signals and events) but not SF-FRP\footcite{Elliott2009}.
    \end{itemize}
\end{frame}

\section{Problem}

\begin{frame}[fragile]
\frametitle{Separating Events and Signals}
    \begin{itemize}
        \item[] 
\begin{Verbatim}
class (Category a) => Arrow a where
  arr :: (b -> c) -> a b c
  ...
\end{Verbatim}
        \item Traditional Signal-Function FRP encodes signal functions as a Haskell Arrow.
\begin{Verbatim}
data SF a b = ...

instance Arrow SF where
  ...
\end{Verbatim}
    \end{itemize}
\end{frame}

\begin{frame}[fragile]
\frametitle{Separating Events and Signals}
    \begin{itemize}
        \item The abstraction must lift any function to a corresponding signal function, without input/output type annotation.
        \item Events encoded as optional signals:
\begin{Verbatim}
type Event a = Maybe a
\end{Verbatim}
        \item Combined signals and events encoded as one signal with\\
              a pair type:
    \end{itemize}
\end{frame}

\begin{frame}[fragile]
\frametitle{Separating Events and Signals}
    \begin{itemize}
        \item[]
\begin{Verbatim}
class (Category a) => Arrow a where
  ...
  first :: a b c -> a (b, d) (c, d)
  ...

capture :: SF (a, Event b) (Event a)

ghci> :t first . arr
first . arr
    :: (Arrow a) => (b -> c) -> a (b, d) (c, d)
\end{Verbatim}
    \end{itemize}
\end{frame}

\section{Approach}

\begin{frame}[fragile]
\frametitle{Type Signal Functions with Signal Vectors}

    \begin{itemize}
        \item Described by Sculthorpe\footcite{Sculthorpe2011}.
        \item Describe separation of individual signals and events.
        \item In Haskell (with {\tt -XEmptyDataDecls}):
\begin{Verbatim}
data SVEmpty
data SVSignal a
data SVEvent a
data SVAppend svl svr
type :^: svl svr = SVAppend svl svr
\end{Verbatim}
        \item Would like to use {\tt -XDataKinds} but it's not working well yet.
    \end{itemize}

\end{frame}

\begin{frame}[fragile]
\frametitle{Type Signal Functions with Signal Vectors}

    \begin{itemize}
        \item Two combinator examples:
        \item Lifting a pure function to transform a signal:
\begin{Verbatim}
pureSignal ::     (a -> b)
               -> SF (SVSignal a) (SVSignal b)
\end{Verbatim}
        \item Passing through input on the right:
\begin{Verbatim}
first ::    SF svIn svOut
         -> SF (svIn :^: svRight) (svOut :^: svRight)
\end{Verbatim}
        \item Composing these leads to:
\begin{Verbatim}
ghci> :t first . pureSignal
first . pureSignal
    :: (a -> b) -> SF (SVSignal a :^: sv)
                      (SVSignal b :^: sv)
\end{Verbatim}
    \end{itemize}
\end{frame}

\begin{frame}
\frametitle{Partial Representations of Signal Vectors}
    \begin{itemize}
        \item To evaluate signals and events differently we must represent them separately.
        \item Signal vectors enable this by distinguishing them in the types.
        \item We can construct several representations of a signal vector.
        \begin{itemize}
          \item Represent signal leaves, event leaves, or both.
          \item Represent one leaf, a subset of leaves, or all applicable leaves.
          \item Transform the type at the leaf, or don't.
        \end{itemize}
    \end{itemize}
\end{frame}

\section{Implementation}

\begin{frame}[fragile]
\frametitle{Signal Representation}
    \begin{itemize} 
        \item Represent an entire signal sample\\ (for initializing a signal function at time zero):
\begin{Verbatim}
data SVSample where
    SVSample      ::    a -> SVSample (SVSignal a)
    SVSampleEvent ::    SVSample (SVEvent a)
    SVSampleEmpty ::    SVSample SVEmpty
    SVSampleBoth  ::    SVSample svLeft 
                     -> SVSample svRight
                     -> SVSample (svLeft :^: svRight) 
\end{Verbatim}
    \end{itemize}
\end{frame}

\begin{frame}[fragile]
\frametitle{Signal Representation}
    \begin{itemize} 
        \item Represent a signal delta\\ (replacement values for a subset of signals):
\begin{Verbatim}
data SVDelta where
     SVDelta        ::   a -> SVDelta (SVSignal a)
     SVDeltaNothing ::   SVDelta sv
     SVDeltaBoth    ::   SVDelta svLeft 
                      -> SVDelta svRight
                      -> SVDelta (svLeft :^: svRight) 
\end{Verbatim}
    \end{itemize}
\end{frame}

\begin{frame}[fragile]
\frametitle{Event Representation}
    \begin{itemize} 
        \item Represent an event occurrence:
\begin{Verbatim}
data SVOccurrence where
     SVOccurrence ::   a -> SVOccurrence (SVEvent a)
     SVOccLeft    ::   SVOccurrence svLeft
                    -> SVOccurrence
                         (svLeft :^: svRight)
     SVOccRight   ::   SVOccurrence svRight
                    -> SVOccurrence
                         (svLeft :^: svRight)
\end{Verbatim}
    \end{itemize}
\end{frame}

\begin{frame}[fragile]
\frametitle{Signal Function Representation}
    \begin{itemize}
        \item Separate continuations for time advancement and event occurrences:
\begin{Verbatim}
data Initialized
data NonInitialized

data SF init svIn svOut where
  SF ::   (SVSample svIn
           -> (SVSample svOut,
               SF Initialized svIn svOut))
       -> SF NonInitialized svIn svOut
...
\end{Verbatim}
    \end{itemize}
\end{frame}

\begin{frame}[fragile]
\frametitle{Signal Function Representation}
    \begin{itemize}
    \item Separate continuations for time advancement and event occurrences.
\begin{Verbatim}
...
  SFInit ::   (Double 
               -> SVDelta svIn
               -> (SVDelta svOut,
                   [SVOccurrence svOut],
                   SF Initialized svIn svOut))
           -> (SVOccurrence svIn
               -> ([SVOccurrence svOut],
                   SF Initialized svIn svOut)
           -> SF Initialized svIn svOut
\end{Verbatim}
    \end{itemize}
\end{frame}

\begin{frame}
\frametitle{Evaluation}
    \begin{itemize}
        \item Yampa/AFRP: Supply SF and input/output actions to an evaluation loop ({\tt reactimate}).
        \item Here: initialize an evaluation state with:
        \begin{itemize}
            \item A signal function.
            \item Initial values for all input signals.
            \item Handlers for all outputs.
        \end{itemize}
        \item Then, the evaluation monad carries this state and provides the actions:
            \begin{itemize}
                \item {\tt push} Push an event (which will be immediately reacted to).
                \item {\tt update} Update the value of an input signal (with no immediate effect).
                \item {\tt step} Update the time and evaluate new values of signals.
            \end{itemize}
    \end{itemize}
\end{frame}

\section{Conclusion}

\begin{frame}
\frametitle{Conclusion}
    \begin{itemize}
        \item Signal vectors fix an abstraction leak in Signal-Function FRP.
        \item Representing events distinctly from signals in the type permits
              heterogeneous (push-pull) evaluation of signal functions.
    \end{itemize}
\end{frame}

\begin{frame}
\frametitle{Further Work}
    \begin{itemize}
        \item Dynamic collections: Dynamically switch between collections of signal functions.
        \item Demonstrations and performance comparisons.
        \item Semantics/correctness proof (especially for event merging).
        \item Time-independence optimization.
    \end{itemize}
\end{frame}

\begin{frame}
\frametitle{Questions?}
\end{frame}

\section{In Detail}

\begin{frame}
\frametitle{Time-independence optimization}
    \begin{itemize}
        \item Inefficiency: running every time continuation every sample step.
        \item This happens even if the signal delta is empty (no change)
        \item Currently this is necessary because the {\em time} delta might
              cause some output.
        \item Optimization:
        \begin{itemize}
            \item Mark time-independent SFs using a separate constructor.
            \item "Smart" composition-routing to make a composite SF time-independent
                  if all of its components are.
            \item Time-independent time continuation has no time input, only delta.
            \item Only called if non-empty delta. 
        \end{itemize}
    \end{itemize}
\end{frame}

\begin{frame}
\frametitle{Event Merging (Semantics Idea)}
    \begin{itemize}
        \item Push-based events are not tied to samples.
        \item We thus know only what the last sample time was for an event,
              not the actual time.
        \item Can we hide event merging in this loss of precision?
    \end{itemize}
\end{frame}

\end{document}