### 2012-10-29 Docs/Thesis/implementation.tex

\chapter{Implementation}
\label{chapter:Implementation}

Having established the design and semantics of the reactive system, I present a purely 
functional implementation. For this implementation, I will make significant use
of extensions to the Haskell type system. These extensions provide mechanisms for expressing
abstractions not easily expressible in a standard Hindley-Milner type system, while maintaining
the type soundness of Haskell.

\input{implementation-signalfunctions}

\input{implementation-evalinterface}