### 2012-10-29 Docs/Thesis/introduction.tex

\chapter{Introduction}
\label{chapter:Introduction}

Most of the useful programs which programmers are asked to write must react to inputs which are not available at the time the program starts,
and produce effects at many different times throughout the execution of the program. Examples of these programs include GUI applications,
web browsers, web applications, video games, multimedia authoring and playback tools, operating system shells and kernels, servers,
robotic control programs, and many others. This attribute of a program is called
reactivity.

Functional reactive programming (FRP) is a promising class of abstractions for writing reactive programs in functional languages. FRP 
abstractions translate desirable properties of the underlying functional languages, such as higher-order functions and referential
transparency, to reactive constructs, generally without modifications to the underlying language. This permits the compositional
construction of reactive programs in purely functional languages.

Functional reactive programming was introduced, though not quite by name, with Fran~\cite{Elliott1997}, a compositional system for writing
reactive animations. From the start, two key challenges emerged: the enforcement of {\em causality}\footnote{Causality is the property 
that a value in an FRP program depends only on present or past values. Obviously, a program which violates this property is impossible to
evaluate, but in some FRP systems such programs are unfortunately quite easy to write down.} in FRP programs, and the efficient
evaluation of FRP programs.

The first challenge was addressed by representing FRP programs, not as
compositions of signals and events, but as compositions of transformers of
behaviors and events called {\em signal functions}. By not permitting direct
access to behaviors or events, but representing them implicitly, signal
functions prohibit accessing past or future time values, only allowing access to
values in the closure of the signal function and current input values. Signal
function FRP programs are written by composing signal functions, and are
evaluated using a function which provides input to the signal function and acts
on the output values. This evaluation function is elided in existing literature
on signal functions, but we provide a documented and improved interface for
evaluating signal functions. A further advantage of signal function FRP programs
is that they are more composable, since additional signal functions may be
composed on the input or output side of a signal function, rather than simply
the output side in classic (non-signal-function) FRP systems.

The second challenge was addressed for classic FRP by the creation of a hybrid
push-pull FRP system~\cite{Elliott2009}. This system relied on the distinction
between reactivity and time dependence to decompose behaviors into phases of
constant or simply time-dependent values, with new phases initiated by event
occurrences. However, this implementation did not address concerns of causality
or compositionality. Further, the existence of events as first-class values in
classic FRP forced the use of an impure and resource-leaking technique to
compare events when merging, to determine which had the first occurrence after
the current time. This technique made use of the Haskell {\tt unsafePerformIO}
function to fork two threads to compare the improving values of the next
occurrence time for two events being merged. Further, since the classic FRP
interface permits access to only the output of a behavior or event, and is
always bound to its implicit inputs, the best the push-based implementation
could do is to block when no computation needs to be performed. The computation
cannot be suspended entirely as a value representation.

To address both challenges in a FRP implementation, we to combine the approaches
and demonstrate a push-based signal-function FRP system. The current
implementation approach for signal-function FRP is inherently
pull-based~\cite{Nilsson2002}, but recent work on N-ary
FRP~\cite{Sculthorpe2011}, a variant of signal-function FRP which enumerates
inputs and distinguishes between events and behaviors, provides a representation
which suggests a method of push-based evaluation. In previous implementations of
FRP, events were simply option-valued signals. This approach has the advantage
of being simple to implement in a pull-based setting. However, it does not have
an obvious semantics for event merging. Further, it necessitates pull-based
evaluation, since there is no way to separate the events which may be pushed
from the truly continuous values which must be repeatedly sampled.

This thesis describes the design and implementation of TimeFlies, a push-pull signal-function FRP system. This system is presented as a library
of combinators for constructing and evaluating FRP programs.