### 2012-10-30 Docs/Thesis/implementation.tex

\chapter{Implementation}
\label{chapter:Implementation}

Having established the design and semantics of the reactive system, I present a purely 
functional implementation. For this implementation, I will make significant use
of an extension to the Haskell type system. This extension, called
GADTs~\cite{Cheney2003,Xi2003}, provides a mechanism for expressing abstractions
which are not easily expressible in a standard Hindley-Milner type system, while
maintaining the type soundness of Haskell. In particular, GADTs provide the
power we require to take advantage of the signal vector representation of signal
function inputs and outputs.

\input{implementation-signalfunctions}

\input{implementation-evalinterface}