### 2012-11-26 Docs/IFL2012/paper.tex
Copypasta thesis chapters into IFL paper. To edit later.

\documentclass[draft]{llncs}

\usepackage{fancyvrb}                                       % For code environment
\DefineVerbatimEnvironment{code}{Verbatim}{fontsize=\small} %

\title{Push-Pull Signal-Function Functional Reactive Programming}
\titlerunning{Push-Pull Signal-Function FRP}

\author{Edward Amsden}
\institute{Rochester Institute of Technology\\\email{eca7215@cs.rit.edu}}

\begin{document}
\maketitle

\begin{abstract}
Functional Reactive Programming is a promising class of systems for writing
interactive and time-dependent programs. Signal-function FRP is a subclass of
these systems which provides advantages of modularity and correctness, but
has proven difficult to efficiently implement.

The abstraction of signal vectors provides the necessary type apparatus to
distinguish components of the input and output of signal functions which benefit
from a push-based implementation from those which benefit from a pull-based
implementation, and to combine both implementation strategies in a single system.

We describe a signal-function FRP system which provides push-based evaluation
for events, pull-based evaluation for signals, and a simple monadic evaluation
interface which permits the system to be easily integrated with one or more
IO systems.
\end{abstract}

\section{Introduction}
\label{section:Introduction}

Functional Reactive Programming (FRP) is a class of systems for describing
reactive programs. Reactive programs are programs which, rather than taking
a single input and producing a single output, must accept multiple inputs and 
alter temporal behavior, including the production of multiple outputs, based
on these inputs.

An FRP system will provide a means of manipulating {\em behaviors} and
{\em events}. Behaviors are often referred to as {\em signals} in FRP literature,
but the definition is the same. A behavior or signal is, semantically, a
function from time to a value. An event is a discrete, possibly infinite, and 
time-ordered sequence of occurrences, which are times paired with values.

FRP systems can generally be categorized as ``classic FRP,'' 
which corresponds to the originally described FRP system in that behaviors
and events are manipulated directly and are first-class values in the FRP
system, or ``signal-function FRP,'' in which behaviors (generally termed
signals in this approach) and events are not first-class values, but signal
functions are first class values. Signal functions are time-dependent and
reactive transformers of signals, events, or combinations of signals and events.

FRP combines behaviors and events through the use of {\em switching}, in which
a behavior (in classic FRP) or a signal function (in signal-function FRP) is
replaced by a new behavior or signal function carried by an event occurrence.

Classic FRP was first described as an system for interactive animations~\cite{Elliott1997}.
Recent work on classic FRP has focused on efficient implementation. One approach to
efficiency is to separate the evaluation of behaviors and events, since suitable 
strategies give best performance in each case. Push-based evaluation evaluates a
system only when input is available, and is thus suitable for discrete inputs
such as events. Pull-based evaluation evaluates the system as quickly as
possible, polling for input, and is preferable for behaviors and signals.
The initial implementations of FRP made use of pull-based evaluation for both
behaviors and events. Reactive~\cite{Elliott2009}, as well as more recent systems such as
``reactive-banana''~\cite{Apfelmus}, make use of push-based evaluation for
events and pull-based evaluation for behaviors. This is known as ``push-pull''
evaluation.

All implementations of signal-function FRP to date~\cite{Courtney2001-1,Nilsson2002,Nilsson2005,Sculthorpe2011}
have used pull-based evaluation for both signals and events. This is due to
the ease of implementation of pull-based evaluation, and the types used for
signal functions which do not permit distinguishing signals and events, or
constructing only part of the input (for instance, one event occurrence.)

A recent extension of signal-function FRP called N-Ary FRP~\cite{Sculthorpe2011}
describes a method of typing signal functions which, as we will show, enables
the push-based evaluation of events in a signal-function FRP system. The notion
of signal vectors allows the representation of signal function inputs and
outputs as combinations of signals and events, rather than a single signal which
may contain multiple values, including option values for events. Signal vectors
are uninhabited types, which can be used to type partial or full representations
of the signal function inputs and outputs.

We present TimeFlies,\footnote{The sentence ``Time flies like an arrow.'' is a 
favorite quotation of one of the author's philosophy instructors, used to
demonstrate the ambiguity of language. The origin of the quotation is unknown.}
a push-pull signal-function FRP system. We hope to demonstrate the feasibility
of such an approach to FRP, and provide a basis for further research into
efficient implementation of signal-function FRP. We also describe a powerful
evaluation interface for TimeFlies, which permits us to use TimeFlies to
describe applications which make use of multiple and differing IO libraries.

Section~\ref{section:System_Design} describes design choices for the system,
and provides an overview of the interface. Section~\ref{section:Implementation}
describes how the system is implemented, and how the separation of evaluation
between events and signals is achieved. Section~\ref{section:Discussion} is a
discussion of the usefulness of our implementation. 
Section~\ref{section:Ongoing_and_Further_Work} describes the current and future
work on this system. Section~\ref{section:Related_Work} gives an overview of
related efforts. Section~\ref{section:Conclusion} concludes.

\section{System Design and Interface}
\label{section:System_Design_and_Interface}

\subsection{Goals}
\label{section:System_Design_and_Interface-Goals}

The goal of FRP is to provide an efficient, declarative abstraction for creating
reactive programs. Towards this overall goal, there are three goals which this
system is intended to meet.

Efficient evaluation is the motivation for push-based evaluation of events.
Since FRP programs are expected to  interact with an outside world in real time,
efficiency cannot simply be measured by the runtime of a program. Thus, when speaking of efficiency,
we are expressing a desire that the system utilize as few system resources as possible
for the task at hand, while responding as quickly as possible to external inputs and
producing output at a consistently high sample rate.

A composable abstraction is one in which values in that abstraction may be
combined in such a way that reasoning about their combined actions involves
little more than reasoning about their individual actions. In a signal function
system, the only interaction between composed signal functions ought to be that
the output of one is the input of another. Composability permits a particularly
attractive form of software engineering in which successively larger systems are
created from by combining smaller systems, without having to reason about the 
implementation of the components of the systems being combined.

It is fine for a system to be composable with regards to itself, but an FRP
system must interact with the outside world. Since we cannot anticipate every
possible form of input and output that the system will be asked to interact
with, we must interface with Haskell's IO system. In particular, most libraries
for user interaction (e.g. GUI and graphics libraries such as GTK+ and GLUT) and
most libraries for time-dependent IO (e.g. audio and video systems) make use of
the event loop abstraction. In this abstraction, event handlers are registered
with the system, and then a command is issued to run a loop which detects events
and runs the handlers, and uses the results of the handlers to render the
appropriate output.

We would like for the FRP system to be easy to integrate with such IO systems,
while being flexible enough to enable its use with other forms of IO systems,
such as simple imperative systems, threaded systems, or network servers.

\subsection{Semantics}
\label{subsection:System_Design_and_Interface-Semantics}

A rigorous and formal elucidation of the semantics of signal-function FRP remains
unattempted, but there is a sufficient change to the practical semantics of
signal-function FRP between previous signal-function systems and TimeFlies to warrant
some description.

In previous systems such as Yampa, events were understood (and implemented) as option-valued
signals. This approach is undesirable for several reasons. The most pressing reason is
that it prohibits push-based evaluation of events, because events are embedded in the
samples of a signal and must be searched for when sampling.

Another concern is that this approach limits the event rate to the sampling rate.
The rate of sampling should, at some level, not matter to the FRP system. Events
which occur between sampling intervals are never observed by the system. Since
the signal function is never evaluated except at the sampling intervals, no
events can otherwise be observed.

This concern drives the next concern. Events are not instantaneous in this formulation.
If a signal is option valued, the sampling instant must fall within the interval where
there is an event occurrence present for that event to be observed. If events are
instantaneous, the probability of observing an event occurrence is zero.

Therefore, TimeFlies employs the N-Ary FRP type formulation to represent signals and
events as distinct entities in the inputs and outputs of signal functions. This means
we are now free to choose our representation of events, and to separate it from the
representation and evaluation of signals.

This freedom yields the additional ability to make events independent of the sampling
interval altogether. The semantics of event handling in TimeFlies is that an event occurrence
is responded to immediately, and does not wait for the next sampling instant. This allows events
to be instantaneous, and further, allows multiple events to occur within a single sampling interval.

There are two tradeoffs intrinsic to this approach. The first is that events are only partially ordered
temporally. There is no way to guarantee the order of observation of event occurrences occurring in the
same sampling interval. Further, the precise time of an event occurrence cannot be observed, only the 
time of the last sample prior to the occurrence.

In return for giving up total ordering and precise observation of the timing of events, we obtain the
ability to employ push-based evaluation for event occurrences, and the ability to non-deterministically
merge event occurrences. When events being input to a non-deterministic merge have simultaneous occurrences,
we arbitrarily select one to occur first. This does not violate any guarantee about time values, since
they will both have the same time value in either case, and does not violate any guarantee about ordering,
since no guarantee of their order is given.

A formal semantic description of signal function FRP would clarify the consequences of this decision somewhat,
but is outside the scope of this thesis.

\subsection{Types}
\label{subsection:System_Design_and_Interface-Types}

In a strongly and statically typed functional language, types are a key part of
an interface. Types provide a mechanism for describing and ensuring properties
of the interface's components and about the systems created with these
components.

In order to type signal functions, we must be able to describe their input and
output. In most signal function systems, a signal function takes exactly one
input and produces exactly one output. Multiple inputs or outputs are handled
by making the input or output a tuple, and combinators which combine or split
the inputs or outputs of a signal assume this. Events are represented at the
type level as a particular type of signal, and at the value level as an option,
either an event occurrence or not.

This method of typing excludes push-based evaluation at the outset.
It is not possible to construct a ``partial tuple'' nor in general is it
possible to construct only part of any type of value. Push-based evaluation
depends on evaluating only that part of the system which is updated, which means
evaluating only that part of the input which is updated.

In order to permit the construction of partial inputs and outputs, we make use
of signal vectors. Signal vectors are uninhabited types which describe the input
and output of a signal function. Singleton vectors are parameterized over the
type carried by the signal or by event occurrences. The definition of the signal
vector type is shown in Figure~\ref{figure:signal_vector_types}.

Having an uninhabited signal vector type allows us to construct representations
of inputs and outputs which are hidden from the user of the system, and are
designed for partial representations.

\begin{figure}
\begin{code}
data SVEmpty    -- An empty signal vector component,
                -- neither event nor signal
data SVSignal a -- A signal, carrying values of type a
data SVEvent a  -- An event, whose occurrences carry values of type a
data SVAppend svLeft svRight -- The combination of the signal vectors
                             -- svLeft and svRight
\end{code}
\hrule
\caption{Signal vector types.}
\label{figure:signal_vector_types}
\end{figure}

The type constructor for signal functions is shown in
Figure~\ref{figure:signal_function_types}. For the {\tt init} parameter, only
one possible instantiation is shown. The usefulness of this type parameter,
along with another instantation which is hidden from users of the library,
is discussed in the section on implementation of signal functions
(Section~\ref{section:Implementation}).

Signal functions with signal vectors as input and output types form a
Haskell {\tt GArrow}~\cite{Megacz2011}. Specifically, the signal function
type constructor (with the initialization parameter fixed) forms the arrow
type, the {\tt SVAppend} type constructor forms the product type, and the
{\tt SVEmpty} type constructor forms the unit type.

The representation of signal functions is discussed in
Section~\ref{section:Implementation-Signal_Functions}. The type synonyms
{\tt :\textasciitilde>} and {\tt :\textasciicircum:} are included for
readability and are not crucial to the FRP system.

\begin{figure}
\begin{code}
-- Signal functions
-- init: The initialization type for 
--   the signal function, always NonInitialized
--   for exported signal functions
-- svIn: The input signal vector
-- svOut: The output signal vector
data SF init svIn svOut

data NonInitialized

type svIn :~> svOut = SF NonInitialized svIn svOut
type svLeft :^: svRight = SVAppend svLeft svRight
\end{code}
\hrule
\caption{Signal function types.}
\label{figure:signal_function_types}
\end{figure}

A {\em monad} is a standard, composable abstraction for writing functions with
a context, used in Haskell for IO~\cite{PeytonJones1993,PeytonJones2001} and
other tasks. A monad is simply a 1-arity type constructor together with two
functions. The first function, {\tt return}, takes a value of type {\tt a} and
produces a value of type {\tt m a}. The second, called {\tt bind} and stylized
in the Haskell standard library as the infix operator {\tt (>>=)}, takes a value
of type {\tt m a} and a function of type {\tt a -> m b} and produces a value of
type {\tt m b}. This allows a value to be operated on out of the context and a
new context to be assigned.

A monad can have other primitives which manipulate the context in some way. For 
instance, the primtives in Haskell's {\tt IO} monad produce actions which, when
interpreted as part of the {\tt main} action, produce some side-effect. The
{\tt State} monad provides {\tt get} and {\tt put} operations to work with a 
state value stored in the context.

Monad transformers~\cite{Jones1995} provide a means to combine the functionality
of multiple monads. A monad transformer is a monad with an extra type parameter.
This type parameter is instantiated with the type constructor of the underlying
monad, and an extra operation {\tt lift} is provided which converts values in
the underlying monad to values in the monad transformer.

The evaluation monad is a monad transformer. This permits it to be used in
conjunction with the {\tt IO} monad (or any other monad) to describe how input
is to be obtained for the signal function being evaluated, and how outputs are
to be handled.

The evaluation monad, in addition to the standard monad operators, provides a
means of {\em initializing} a signal function, and a means of translating the
monadic value describing evaluation to a value in the underlying monad. This
means, for instance, that we can obtain an action in the {\tt IO} monad to
evaluate a signal function.

The type of the evaluation monad must track the input and output type of the
signal function. The monad's context stores a mapping from outputs to handling
actions. There are thus four type parameters to the monad's type
constructor: the input signal vector, the output signal vector, the type of the
underlying monad, and the monadic type parameter. The type is shown in
Figure~\ref{figure:evaluation_monad_types}.

\begin{figure}
\begin{code}
-- A signal function's evaluations state
data SFEvalState svIn svOut m
-- The evaluation monad
data SFEvalT svIn svOut m a

instance Monad m => Monad (SFEvalT svIn svOut m)
\end{code}
\hrule
\caption{Evaluation monad types.}
\label{figure:evaluation_monad_types}
\end{figure}

\subsection{Combinators}
\label{section:System_Design_and_Interface-Combinators}

Signal functions are constructed from combinators, which are primitive signal
functions and operations to combine these primitives. These combinators are
grouped as basic signal functions, lifting operations for pure functions,
routing, reactivity, feedback, event processing, joining, and time dependence.

The basic signal functions (Figure~\ref{figure:basic_signal_functions})
provide very simple operations. The {\tt identity} signal function, as expected,
simply copies its input to its output. The {\tt constant} signal function
produces the provided value as a signal at all times. The {\tt never} signal
function has an event output which never produces occurrences. The {\tt asap}
function produces an event occurrence with the given value at the first time
step after it is switched into the network. The {\tt after} function waits for
the specified amount of time before producing an event occurrence.

With the exception of {\tt identity}, all of the basic signal functions have
empty inputs. This allows these signal functions to be used to insert values
into the network which are known when the signal function is created, without
having to route those values from an input.

\begin{figure}
\begin{code}
-- Pass the input unmodified to the output
identity :: sv :~> sv

-- Produce a signal which is at all times the supplied value
constant :: a -> SVEmpty :~> SVSignal a

-- An event with no occurrences
never    :: SVEmpty :~> SVEvent a

-- An event with one occurrence, as soon as possible after
-- the signal function is initialized
asap     :: a -> SVEmpty :~> SVEvent a

-- An event with one occurrence
-- after the specified amount of time has elapsed.
after    :: Double -> a -> SVEmpty :~> SVEvent a
\end{code}
\hrule
\caption{Basic signal function combinators.}
\label{figure:basic_signal_functions}
\end{figure}

Two combinators are provided to lift pure functions to signal functions
(Figure~\ref{figure:lifting_pure_functions}). The {\tt pureSignalTransformer}
combinator applies the pure function to a signal at every sample point. The
{\tt pureEventTransformer} combinator applies the function to each occurrence of
an input event.

\begin{figure}
\begin{code}
-- Apply the given function to a signal at all points in time
pureSignalTransformer :: (a -> b) -> SVSignal a :~> SVSignal b

-- Apply the given function to each event occurrence
pureEventTransformer  :: (a -> b) -> SVEvent a :~> SVEvent b
\end{code}
\hrule
\caption{Lifting pure functions.}
\label{figure:lifting_pure_functions}
\end{figure}

The routing combinators are used to combine signal functions, and
to re-arrange signal vectors in order to connect signal functions.
The routing combinators are shown in Figure~\ref{figure:routing_combinators}.

Only those combinators which modify or combine signal functions
({\tt (>>>)}, {\tt first}, {\tt second}) are reactive (may replace themselves in
response to events, and then only if they inherit their reactivity from the
signal function(s) which are inputs to combinator. The rest do not
react to or modify the input in any way, except to re-arrange it, copy it, or
discard it altogether.

\begin{figure}
\begin{code}
-- Use the output of one signal function as the input for another
(>>>) :: (svIn :~> svBetween) -> (svBetween :~> svOut) -> svIn :~> svOut

-- Pass through the right side of the input unchanged
first :: (svIn :~> svOut) -> (svIn :^: sv) :~> (svOut :^: sv)

-- Pass through the left side of the input unchanged
second :: (svIn :~> svOut) -> (sv :^: svIn) :~> (sv :^: svOut)

-- Swap the left and right sides
swap :: (svLeft :^: svRight) :~> (svRight :^: svLeft)

-- Duplicate the input
copy :: sv :~> (sv :^: sv)

-- Ignore the input
ignore :: sv :~> svEmpty

-- Remove an empty vector on the left
cancelLeft :: (SVEmpty :^: sv) :~> sv

-- Add an empty vector on the left
uncancelLeft :: sv :~> (SVEmpty :^: sv)

-- Remove an empty vector on the right
cancelRight :: (sv :^: SVEmpty) :~> sv

-- Add an empty vector on the right
uncancelRight :: sv :~> (sv :^: SVEmpty)

-- Make right-associative
associate :: ((sv1 :^: sv2) :^: sv3) :~> (sv1 :^: (sv2 :^: sv3))

-- Make left-associative
unassociate :: (sv1 :^: (sv2 :^: sv3)) :~> ((sv1 :^: sv2) :^: sv3)
\end{code}
\hrule
\caption{Routing combinators.}
\label{figure:routing_combinators}
\end{figure}

Reactivity is introduced by means of the {\tt switch} combinator
(Figure~\ref{figure:switch_combinator}). The design of this combinator
allows modular packaging of reactivity. A signal function can determine
autonomously when to replace itself, based only on its input and state,
by emitting an event occurrence carrying its replacement. The combinator
consumes and hides the event carrying the replacement signal function, 
so the reactivity is not exposed by the resulting reactive signal function.

\begin{figure}
\begin{code}
switch ::    (svIn :~> (svOut :^: SVEvent (svIn :~> svOut)))
          -> svIn :~> svOut
\end{code}
\hrule
\caption{Reactivity combinator.}
\label{figure:switch_combinator}
\end{figure}

There are other formulations of a reactive combinator found in FRP literature
and libraries, which may be implemented using the one supplied. These are shown
in Figure~\ref{figure:alternate_switching_combinators} and may be provided in a
future version of the TimeFlies library.

\begin{figure}
\begin{code}
-- Alternate version of switch,
-- implemented in terms of supplied version
switch_gen ::    (svIn :~> (svOut :^: SVEvent a))
              -> (a -> svIn :~> svOut)
              -> svIn :~> svOut
switch_gen sf f =
  switch (sf >>> second (pureEventTransformer f))

-- Supplied version in terms of alternate version
switch ::    (svIn :~> (svOut :^: SVEvent (svIn :~> svOut)))
          -> svIn :~> svOut
switch sf = switch_gen sf id

-- Repeated switch, which takes replacement signal functions
-- externally.
rswitch ::    (svIn :~> svOut)
           -> (svIn :^: SVEvent (svIn :~> svOut)) :~> svOut
rswitch sf =
  switch (first sf >>> second (pureEventTransformer rswitch))
\end{code}
\hrule
\caption{Alternate reactivity combinators.}
\label{figure:alternate_switching_combinators}
\end{figure}

It is often useful for a signal function to receive a portion of its
own output as input. This is especially useful when we have two
signal functions which we would like to mutually interact. We cannot
just serially compose them, we must also bring the output of the second
back around to the first. Many signal-processing algorithms also depend
on feedback. The combinator which provides this ability is shown
in Figure~\ref{figure:feedback_combinator}.

\begin{figure}
\begin{code}
loop ::    ((svIn :^: svLoop) :~> (svOut :^: svLoop))
        -> svIn :~> svOut
\end{code}
\hrule
\caption{Feedback combinator.}
\label{figure:feedback_combinator}
\end{figure}

This combinator provides decoupling for signals
(the input signal is the output signal at the previous time-step)
but not events (event occurrences are supplied to the combinator immediately).
This means that the programmer has the responsibility to ensure that feedback
does not generate an infinite sequence of events in a single time-step.

Several combinators are provided for manipulating, suppressing, and generating events.
Each of the combinators has an option variant and a list variant. The option variant
produces an output event occurrence whenever the application of the supplied function
to the input event produces a value. The list version produces an event occurrence for
each of the elements of the output list. The combinators are shown in Figure~\ref{figure:event_specific_combinators}.

\begin{figure}
\begin{code}
-- Apply the function to each input occurrence,
-- and produce an occurrence for each Just.
filter :: (a -> Maybe b) -> SVEvent a :~> SVEvent b

-- Apply the function to each input occurrence,
-- and produce an occurrence for each list element
filterList :: (a -> [b]) -> SVEvent a :~> SVEvent b

-- Apply the function to the stored accumulator
-- and the event occurrence, replacing the accumulator
-- and possibly outputting an occurrence
accumulate ::    (a -> b -> (Maybe c, a))
              -> a
              -> SVEvent b :~> SVEvent c

-- Apply the function to the stored accumulator
-- and the event occurrence, replacing the
-- accumulator and outputting an event occurrence
-- for each element of the list
accumulateList ::    (a -> b -> ([c], a))
                  -> a
                  -> SVEvent b :~> SVEvent c
\end{code}
\hrule
\caption{Event-specific combinators.}
\label{figure:event_specific_combinators}
\end{figure}

The filter combinators are stateless, and thus apply the function to only the new
input value. They are useful for suppressing events, as well as for extracting one
of multiple cases of a datatype. For instance, a splitter for events carrying
{\tt Either}-valued occurrences could be written as:

\begin{code}
getLeft :: Either a b -> Maybe a
getLeft (Left x) = Just x
getLeft _ = Nothing

getRight :: Either a b -> Maybe b
getRight (Right x) = Just x
getRight _ = Nothing

split :: SVEvent (Either a b) :~> (SVEvent a :^: SVEvent b)
split = copy >>> first (filter getLeft) >>> second (filter getRight)
\end{code}

The accumulate combinators are stateful, applying the supplied function
to both the input value and an accumulator. This function has two results:
the option or list of output event occurrence values, and the new value
for the accumulator.

The accumulator is useful when responses to multiple event occurrences
(from one or more sources) must be coordinated. For instance, in the
benchmark application (see Chapter~\ref{chapter:Evaluation_and_Comparisons}),
a table is maintained that allows knowledge from previous event occurrences
(packets from a network switch) to be used in deciding where the present
packet ought to go.

The joining combinators provide the ability to combine two event
streams, two signals, or a signal and an event stream. These
combinators are shown in Figure~\ref{figure:joining_combinators}

The {\tt union} combinator is a non-deterministic merge of event
streams. Any event which occurs on either input will occur
on the output. For simultaneous event occurrences, the order of occurrence
is not guaranteed, but the occurrence itself is. This construct
is also guaranteed to respect the relationship of event occurrences to sampling
intervals.

The {\tt combineSignals} combinator applies a binary function pointwise to two
signals, and produces the result of this application as a third signal. The
combining function is necessary because we will have two input samples at each
time step, and must produce exactly one output sample.

The {\tt capture} combinator adds the last-sampled value of a signal at the time
of an event occurrence to that event occurrence.

These three combinators together provide the ability to combine elements of
signal vectors. By combining these combinators, along with the {\tt cancelLeft}
and {\tt cancelRight} routing combinators, arbitrary signal vectors can be
reduced.

\begin{figure}
\begin{code}
union          :: (SVEvent a :^: SVEvent a) :~> SVEvent a
combineSignals :: (a -> b -> c) -> (SVSignal a :^: SVSignal b) :~> SVSignal c
capture        :: (SVSignal a :^: SVEvent b) :~> SVEvent (b, a)
\end{code}
\hrule
\caption{Joining combinators.}
\label{figure:joining_combinators}
\end{figure}

A set of combinators are provided for making use of time-dependence in a signal
function. These combinators allow the modification of signals and events with
respect to time, and the observation of the current time value. The combinators
are shown in Figure~\ref{figure:time_combinators}

\begin{figure}
\begin{code}
time      :: SVEmpty :~> SVSignal Double
delay     :: Double -> (SVEvent a :^: SVEvent Double) :~> SVEvent a
integrate :: TimeIntegrate i => SVSignal i :~> SVSignal i
\end{code}
\hrule
\caption{Time-dependent combinators.}
\label{figure:time_combinators}
\end{figure}

The simplest time-dependent combinator is {\tt time}, which simply outputs
the time since it began to be evaluated. This does not necessarily correspond to
the time since the global signal function began to be evaluated, since the
signal function in which the {\tt time} combinator is used may have been
introduced through {\tt switch}.

The {\tt delay} signal function allows events to be delayed. An initial delay
time is given, and event occurrences on the right input can carry a new delay
time. Event occurrences on the left input are stored and output when their delay
time has passed. Changing the delay time does not affect occurrences already
waiting.

The {\tt integrate} combinator outputs the rectangle-rule approximation of the
integral of its input signal with respect to time.

\subsection{Evaluation Interface}
\label{subsection:System_Design_and_Interface-Evaluation_Interface}

The evaluation interface provides a modified state monad which holds a signal
function, together with some additional information, as its state (shown in Figure~\ref{figure:evaluation_state}.
Rather than monadic instructions to put and get the state, the monad provides instructions
to trigger an input event, update an input signal, and trigger sampling of
signals in the signal function. Additional state includes the current set of
modifications to the input signals (since the last sample) and a set of
handlers which actuate effects based on output events or changes to the output
signal.

\begin{figure}
\begin{code}
-- A vector of handlers for outputs
data SVHandler out sv

-- A dummy handler for an empty output
emptyHandler    :: SVHandler out SVEmpty

-- A handler for an updated signal sample
signalHandler   :: (a -> out) -> SVHandler out (SVSignal a)

-- A handler for an event occurrence
eventHandler    :: (a -> out) -> SVHandler out (SVEvent a)

-- Combine handlers for a vector
combineHandlers ::    SVHandler out svLeft
                   -> SVHandler out svRight
                   -> SVHandler out (svLeft :^: svRight)

-- The state maintained when evaluating a signal function
data SFEvalState m svIn svOut

-- Create the initial state for evaluating a signal function
initSFEval ::    SVHandler (m ()) svOut   -- Output handlers

              -> SVSample svIn            -- Initial input samples

              -> Double                   -- Initial external time,
                                          -- corresponding to time 0 for
                                          -- the signal function

              -> (svIn :~> svOut)         -- Signal function to evaluate
              -> SFEvalState m svIn svOut
\end{code}
\hrule
\caption{State maintained when evaluating a signal function.}
\label{figure:evaluation_state}
\end{figure}

Sets which correspond to signal vectors are built with ``smart'' constructors.
For instance, to construct a set of handlers, individual handling functions are
lifted to handler sets with the {\tt signalHandler} and {\tt eventHandler}
functions, and then combined with each other and {\tt emptyHandler} leaves
using the {\tt combineHandlers} function.

Building the initial input sample is similar, but {\tt sampleEvt} leaves do
not carry a value.

In order to initialize the state, the user must supply a set of handlers, the
signal function to evaluate, and initial values for all of the signal inputs
(Figure~\ref{figure:initial_input}).

\begin{figure}
\begin{code}
-- A sample for all leaves of a signal vector
data SVSample sv

-- Create a sample for a signal leaf
sample          :: a -> SVSample (SVSignal a)

-- A dummy sample for an event leaf
sampleEvt       :: SVSample (SVEvent a)

-- A dummy sample for an empty leaf
sampleNothing   :: SVSample SVEmpty

-- Combine two samples
combineSamples  ::    SVSample svLeft
                   -> SVSample svRight
                   -> SVSample (svLeft :^: svRight)
\end{code}
\hrule
\caption{Data type for initial input}
\label{figure:initial_input}
\end{figure}

This state can then be passed to a monadic action which will supply input to
the signal function. Inputs are constructed using a simple interface with
functions to construct sample updates and event occurrences, and to specify
their place in the vector (Figure~\ref{figure:ongoing_input}).

\begin{figure}
\begin{code}
-- Class to overload left and right functions
class SVRoutable r where
  svLeft          :: r svLeft -> r (svLeft :^: svRight)
  svRight         :: r svRight -> r (svLeft :^: svRight)

-- An input event occurrence
data SVEventInput sv
instance SVRoutable SVEventInput sv

-- An updated sample for a signal
data SVSignalUpdate sv
instance SVRoutable SVSignalUpdate sv

-- Create an event occurrence
svOcc           :: a -> SVEventInput (SVEvent a)

-- Create an updated sample
svSig           :: a -> SVSignalUpdate (SVSignal a)
\end{code}
\hrule
\caption{Data types for ongoing input.}
\label{figure:ongoing_input}
\end{figure}

The {\tt SFEvalT} monad is actually a monad transformer, that is, it is
parameterized over an underlying monad whose actions may be lifted to
{\tt SFEvalT}. In the usual case, this will be the {\tt IO} monad.

{\tt SFEvalT} actions are constructed using combinators to push events,
update inputs, and step time, as well as actions lifted from the underlying
monad (used to obtain these inputs). An action in the underlying monad
which produces the result and a new state is obtained with the {\tt runSFEvalT}
function. These combinators are shown in Figure~\ref{figure:evaluation_combinators}.

\begin{figure}
\begin{code}
-- The evaluation monad
data SFEvalT svIn svOut m a
instance MonadTrans (SFEvalT svIn svOut)
instance (Monad m) => Monad (SFEvalT svIn svOut m)
instance (Functor m) => Functor (SFEvalT svIn svOut m)
instance (Monad m, Functor m) => Applicative (SFEvalT svIn svOut m)
instance (MonadIO m) => MonadIO (SVEvalT svIn svOut m)

-- Obtain an action in the underlying monad
-- from an SFEvalT and a new state.
runSFEvalT ::    SFEvalT svIn svOut m a
              -> SFEvalState m svIn svOut
              -> m (a, SFEvalState m svIn svOut)

-- Push an event occurrence.
push :: (Monad m) => SVEventInput svIn -> SFEvalT svIn svOut m ()

-- Update the value of an input signal sample
-- (not immediately observed)
update :: (Monad m) => SVEventInput svIn -> SFEvalT svIn svOut m ()

-- Step forward in time, observing the updated signal values
step :: (Monad m) => Double -> SFEvalT svIn svOut m ()
\end{code}
\hrule
\caption{Evaluation combinators}
\label{figure:evaluation_combinators}
\end{figure}

\section{Implementation}
\label{section:Implementation}

We now turn our attention to the implementation of the signal function
system. We will discuss representations of inputs, outputs, and signal functions,
as well as the implementations of specific signal function combinators.

\subsection{Signal Functions}
\label{subsection:Implementation-Signal_Functions}

The design of signal functions specifies a family of types for the inputs and
outputs of signal functions. Signal functions are not functions in the purest
sense, however. They are not mappings from a single instance of their input
type to a single instance of their output type. They must be implemented with
respect to the temporal semantics of their inputs and outputs.

We therefore start by creating a set of concrete datatypes for the inputs and
outputs of signal functions. These datatypes will be parameterized by the input
and output types of the signal function, and will not be exposed to the user of
the library. Rather, they will specify how data is represented during the
temporal evaluation of signal functions.

We then describe how signal functions are implemented using these concrete
types, along with higher-order functions and continuations.

In Section~\ref{section:System_Design_and_Interface} we presented signal vectors
as a set of types. In order to be completely principled, we should isolate these
types into their own {\em kind} (a sort of type of types); however, the Haskell
extension for this was far from stable at the time this system was created.

The types are therefore expressed in the system exactly as they were described
in Section~\ref{section:System_Design_and_Interface}. (To refresh, see
Figure~\ref{figure:signal_vector_types}.) The striking observation about these
types is that they have {\em no data constructors}. There are no values which
take these types.

Instead, we will create concrete representations which are parameterized over
these types. These concrete representations will be expressed as GADTs, allowing
each data constructor of the representation to fill in a specific signal vector
type for the parameter of the representation.

The first thing to represent is {\em samples}, which are sets of values for
the signal components of a signal vector. Therefore, we create a representation
which carries a value for every {\tt SVSignal} leaf of a signal vector. In order
to do this, we restrict each of our constructors to a single signal vector type.
So there are three leaf constructors: {\tt SVSample}, which carries a value; and
{\tt SVSampleEvent} and {\tt SVSampleEmpty}, which do not. This ensures that the
only way to represent a sample leaf is with the {\tt SVSample} constructor,
which carries a value of the appropriate type. The datatype is shown in
Figure~\ref{figure:signal_sample_datatype}.

\begin{figure}
\begin{code}
data SVSample sv where
  SVSample      ::    a
                   -> SVSample (SVSignal a)
  SVSampleEvent ::    SVSample (SVEvent a)
  SVSampleEmpty ::    SVSample SVEmpty
  SVSampleBoth  ::    SVSample svLeft
                   -> SVSample svRight
                   -> SVSample (SVAppend svLeft svRight)
\end{code}
\hrule
\caption{Datatype for signal samples.}
\label{figure:signal_sample_datatype}
\end{figure}

What about the event components? We want to represent event occurrences,
each of which will correspond to at most one event in the vector. So a different
representation is called for. In this case, there will be only three
constructors. One constructor will represent an event leaf, and the other will
represent a single value on the left or right side of the node ({\tt SVAppend}),
ignoring all of the type structure on the other side. This representation
describes a path from the root of the signal vector, terminating at an event
leaf with a value.

By pattern matching on the path constructors, we can determine which subvector
of a signal vector an event occurrence belongs to, repeatedly refining it until
we determine which event in the vector the occurrence corresponds to. The
datatype for occurrences is shown in Figure~\ref{figure:event_occurrence_datatype}.

\begin{figure}
\begin{code}
data SVOccurrence sv where
  SVOccurrence ::    a
                  -> SVOccurrence (SVEvent a)
  SVOccLeft    ::    SVOccurrence svLeft
                  -> SVOccurrence (SVAppend svLeft svRight)
  SVOccRight   ::    SVOccurrence svRight 
                  -> SVOccurrence (SVAppend svLeft svRight)
\end{code}
\hrule
\caption{Datatype for event occurrences.}
\label{figure:event_occurrence_datatype}
\end{figure}

We add one more representation for signals, in order to avoid uneccessary
representations of the values of all signals when not all signals have changed
their values. This representation allows us to represent the values of zero or
more of the signals in a signal vector. To accomplish this, we replace the
individual constructors for the {\tt SVEmpty} and {\tt SVEvent} leaves with %there is a better word, something about "useless but not really", for this
a single, unconstrained constructor. This constructor can represent an arbitrary
signal vector. We can use the constructor for signal vector nodes and the 
constructor for sample leaves to represent the updated values, while filling
in the unchanged portions of the signal vector with this general constructor.
This datatype is shown in Figure~\ref{figure:signal_update_datatype}.

\begin{figure}
\begin{code}
data SVDelta sv where
  SVDeltaSignal  ::    a
                    -> SVDelta (SVSignal a)
  SVDeltaNothing ::    SVDelta sv
  SVDeltaBoth    ::    SVDelta svLeft
                    -> SVDelta svRight
                    -> SVDelta (SVAppend svLeft svRight)
\end{code}
\hrule
\caption{Datatype for signal updates.}
\label{figure:signal_update_datatype}
\end{figure}

We now have concrete datatypes for an implementation to operate on. Our next
task is to represent transformers of temporal data, which themselves may change
with time. The common approach to this task is sampling, in which a program
repeatedly checks for updated information, evaluates it, updates some state,
and produces an output. This is the essence of pull-based evaluation.

Another approach is notification, in which the program exposes an interface
which the source of updated information may invoke. This is a repeated entry
point to the program, which causes the program to perform the same tasks
listed above, namely, evaluate the updated information, update state, and
produce output. The strategy of notification as opposed to repeated checking is
the essence of push-based evaluation.

Signal functions are declarative objects, and not running processes. They have
no way to invoke sampling themselves. They can, however, expose separate
interfaces for when sampling is invoked, and when they are notified of an event
occurrence. This creates two control paths through a signal function. One of
these control paths is intended to be invoked regularly and frequently with
updates to the time and sample values, and the other is intended to be invoked
only when an event occurs. The benefit of separating these control paths is that
events are no longer defined in terms of sampling intervals, and need not even
be considered in sampling, except when they are generated by a condition on a
sample. On the other hand, events can be responded to even if the time has not
yet come for another sample, and multiple events can be responded to in a single
sampling interval.

We represent signal functions as a GADT with three type parameters and two 
constructors. The first type parameter represents the initialization state,
and is specialized to {\tt Initialized} or {\tt NonInitialized} depending on the
constructor. The other two type parameters are the input and output signal
vectors, respectively. The signal functions that a user will compose are\
non-initialized signal functions. They must be provided with an initial set of
input signals (corresponding to time zero). When provided with this input, they
produce their time-zero output, and an initialized signal function. The datatype
is shown in Figure~\ref{figure:signal_function_datatype}.

\begin{figure}
\begin{code}
data Initialized

data NonInitialized

data SF init svIn svOut where
  SF     ::    (SVSample svIn 
                  -> (SVSample svOut,
                      SF Initialized svIn svOut)) 
            -> SF NonInitialized svIn svOut
  SFInit ::    (Double 
                  -> SVDelta svIn
                  -> (SVDelta svOut,
                      [SVOccurrence svOut],
                      SF Initialized svIn svOut)) 
            -> (SVOccurrence svIn
                  -> ([SVOccurrence svOut],
                      SF Initialized svIn svOut))
            -> SF Initialized svIn svOut
\end{code}
\hrule
\caption{Datatype and empty types for signal functions.}
\label{figure:signal_function_datatype}
\end{figure}

Initialized signal functions carry two continuations. The first continuation
takes a time differential and a set of signal updates, and returns a set of
signal updates, a collection of event occurrences, and a new initialized signal
function of the same type. This is the continuation called when sampling.

The second continuation takes an event occurrence, and returns a collection of
event occurrences and a new signal function of the same type. This continuation
is only called when there is an event occurrence to be input to the signal
function.

Note that each of these continuations uses one or more of the concrete
representations of signal vectors, and applies the type constructor for the
representation to the input or output signal vector for the signal function.

Having specified a datatype for signal functions, we must now provide
combinators which produce signal functions of this type. Each combinator's
implementation must specify how it is initialized, how it samples its input, and
how it responds to event occurrences.

We will not detail every combinator here, but we will discuss each of the
implementation challenges encountered.

As an example of the implementation of combinators, we show the implementation
of the {\tt identity} signal function in Figure~\ref{figure:identity_implementation}.
This signal function simply passes all of its inputs along as outputs. The
initialization function simply passes along the received sample and outputs the
initialized version of the signal function. The initialized version of the input
is similar, but is self-referential. It outputs itself as its replacement. This
is standard for simple and routing combinators which are not reactive, and
simply move samples and event occurrences around.

\begin{figure}
\begin{code}
identity :: sv :~> sv
identity =
  SF (\initSample -> (initSample, identityInit))

identityInit :: SF Initialized sv sv
identityInit =
  SFInit (\dt sigDelta -> (sigDelta, [], identityInit))
         (\evtOcc -> ([evtOcc], identityInit))
\end{code}
\hrule
\caption{Implementation of the {\tt identity} combinator.}
\label{figure:identity_implementation}
\end{figure}

In order for our primitive signal functions to be useful, we need a means of
composing them. Serial composition creates one signal function from two, by
using the output of one as the input of the other. The serial composition
combinator is styled {\tt (>>>)}. The implementation of this operator is one
place where the advantage of responding to events independently from signal
samples becomes clear.

This is the only primitive combinator which takes two signal functions, and
thus, it is the only way to combine signal functions. Parallel, branching, and
joining composition can be achieved by modifying signal functions with the
{\tt first} and {\tt second} combinators and composing them with the
routing and joining combinators.

Combinators which take one or more signal functions as input must recursively
apply themselves, as is shown in the implementation of serial composition
(Figure~\ref{figure:serial_composition_implementation}). They must also
handle initialization, retaining the initialized signal functions and passing
them to the initialized version of the combinator.

\begin{figure}
\begin{code}
(>>>) ::    (svIn :~> svBetween) 
         -> (svBetween :~> svOut)
         -> (svIn :~> svOut)
(SF sigSampleF1) >>> (SF sigSampleF2) =
  SF (\sigSample -> let (sigSample', sfInit1) = sigSampleF1 sigSample
                        (sigSample'', sfInit2) = sigSampleF2 sigSample'
                    in (sigSample'', composeInit sfInit1 sfInit2))

composeInit ::     SF Initialized svIn svBetween
                -> SF Initialized svBetween svOut
                -> SF Initialized svIn svOut
composeInit (SFInit dtCont1 inputCont1) sf2@(SFInit dtCont2 inputCont2) =
  SFInit
    (\dt sigDelta -> 
       let (sf1MemOutput, sf1EvtOutputs, sf1New) = dtCont1 dt sigDelta
           (sf2MemOutput, sf2EvtOutputs, sf2New) = dtCont2 dt sf1MemOutput
           (sf2EvtEvtOutputs, sf2Newest) = applySF sf2New sf1EvtOutputs
       in (sf2MemOutput,
           sf2EvtOutputs ++ sf2EvtEvtOutputs,
           composeInit sf1New sf2Newest)
    )
    (\evtOcc -> 
      let (sf1Outputs, newSf1) = inputCont1 evtOcc
          (sf2FoldOutputs, newSf2) = applySF sf2 sf1Outputs
      in (sf2FoldOutputs, composeInit newSf1 newSf2)   
    )

applySF ::    SF Initialized svIn svOut
           -> [SVOccurrence svIn]
           -> ([SVOccurrence svOut],
               SF Initialized svIn svOut)
applySF sf indices =
  foldr (\evtOcc (changes, SFInit _ changeCont) ->
           let (newChanges, nextSF) = changeCont evtOcc
               in (newChanges ++ changes, nextSF))
        ([], sf)
        indices
\end{code}
\hrule
\caption{Implementation of serial composition.}
\label{figure:serial_composition_implementation}
\end{figure}

The switch combinator is the means of introducing reactivity into a signal
function. This combinator allows a signal function to replace itself by
producing an event occurrence. The combinator wraps a signal function, and 
observes an event on the right side of the output signal vector. At the first
occurrence of the event, the signal function carried by the occurrence replaces
the signal function.

The switch combinator stores the input sample provided during initialization,
and updates it with the input signal updates. When the wrapped signal function
produces an occurrence carrying a new signal function, that signal function is
initialized with the stored input sample. It is then "wrapped" by another
function which closes over its output sample, and outputs the sample as a signal
update as the next time step. After this, it acts as the new signal function.
This wrapping has some performance implications, which are discussed

This combinator checks the outputs of the wrapped
signal function for an event occurrence from which an uninitialized signal
function is extracted. The switch combinator stores the full sample
for its input vector (which is identical to the input vector of the supplied
signal function) to initialize the new signal function. This also demands that
it add a wrapper to the new signal function which waits for the next sampling
interval and actuates the sample output at initialization as an output set
of changes to the signal. This has some performance implications, which are
discussed in Chapter~\ref{chapter:Evaluation_and_Comparisons}.

Most of the routing combinators are simple to implement. The only task is to add
remove, or replace routing constructors on signal updates and event occurrences.
Since these signal functions are stateless and primitive, they can simply
return themselves as their replacements. The {\tt swap} combinator is shown as
an example in Figure~\ref{figure:swap_implementation}

\begin{figure}
\begin{code}
-- T.swap is imported from Data.Tuple
swap :: (sv1 :^: sv2) :~> (sv2 :^: sv1)
swap =
  SF ((, swapInit) . 
      uncurry combineSamples .
      T.swap . splitSample)

swapInit :: SF Initialized (SVAppend sv1 sv2) (SVAppend sv2 sv1)
swapInit =
  SFInit (flip (const .
                (, [], swapInit) .
                uncurry combineDeltas .
                T.swap . splitDelta))
          (\evtOcc ->
             (case chooseOccurrence evtOcc of
                Left lOcc  -> [occRight lOcc]
                Right rOcc -> [occLeft rOcc], swapInit))

\end{code}
\hrule
\caption{Implementation of the {\tt swap} routing combinator.}
\label{figure:swap_implementation}
\end{figure}

The {\tt first} and {\tt second} combinators are similar to serial composition,
but they transform only one signal function. For signal changes, they must split
the set of input changes into those which will be passed to the signal function
and those which will be simply passed along to the output, and then recombine
them on the other side. For event occurrences, the occurrence must be
pattern-matched to determine whether to call the event continuation from the
provided signal function or passed through, and output event occurrences must
have the suitable routing constructor re-applied. In any case, when a
continuation has been applied, the combinator must be recursively applied to the
new signal function.

The looping feedback combinator is particularly tricky. As it is currently
implemented, the initial sample for the right side of the input signal vector to
the supplied function is the right side of the output sample. This is acceptable,
given Haskell's non-strict evaluation strategy, but it is necessary that the
right side of the signal function's output not be immediately dependent on its
input. The feedback combinator makes use of Haskell's lazy evaluation to
feed events back into the combinator, and stores signal updates until the next
sample. Signal samples are thus automatically decoupled after initialization.
The implementation makes use of the recursive nature of the {\tt let} construct
in Haskell, and the non-strict evaluation of Haskell, to implement feedback.

The {\tt filter} combinators are simple to implement. Their sampling
continuation is superfluous, and the event continuation merely applies the
supplied function, and constructs an output list based on the result.

The {\tt accumulate} combinators are implemented in terms of the {\tt filter}
and {\tt switch} combinators, as shown in
Figure~\ref{figure:accumulate_implementation}.

\begin{figure}
\begin{code}
-- | Accumulate over event occurrences
accumulate :: (a -> b -> (Maybe c, a)) -> a -> SVEvent b :~> SVEvent c
accumulate f a = acc a
  where acc a = switch (pureEventTransformer (f a) >>>
                        copy >>>
                        first (pureEventTransformer fst >>> filter id) >>>
                        second (pureEventTransformer (acc . snd)))

-- | Accumulate over event occurrences, with lists of event outputs
accumulateList :: (a -> b -> ([c], a)) -> a -> SVEvent b :~> SVEvent c
accumulateList f a = acc a
  where acc a = switch (pureEventTransformer (f a) >>>
                        copy >>>
                        first (pureEventTransformer fst >>> filterList id) >>>
                        second (pureEventTransformer (acc . snd)))
\end{code}
\hrule
\caption{Implementation of event accumulators.}
\label{figure:accumulate_implementation}
\end{figure}

The implementation of the joining combinators is simple. The {\tt union}
combinator simply passes along every event occurrence it receives on either
input, stripping off the left and right combinator. This is acceptable since we
do not insist on a total ordering of events, or an event time resolution greater
than the sampling rate. The {\tt combineSignals} combinator maintains the values
of both signals, and applies the combination function whenever one is updated.
The {\tt capture} combinator maintains the input signal value, and adds it to
each event occurrence.

Time dependence is introduced by the {\tt time}, {\tt delay}, and {\tt integrate}
combinators. The time combinator simply sums the time updates and provides the
sum as a signal output. The {\tt delay} combinator keeps a table of events
which have come in, along with their schedule occurrence time, and produces
them as output when time advances far enough. The integrate combinator performs
rectangle-rule integration on signal samples with respect to time.

The implementation strategy leaves room for optimizations. In particular, an
additional constructor for time-independent signal functions would allow
portions of a signal function to forgo evaluation during time steps unless they
had signal updates. Optimizations in the style of Yampa, observed by keeping
an updated AST for the signal function and pattern-matching on it when switching,
might further improve performance. In particular, collapsing nested or
serially-composed versions of the {\tt switchWait} step when switching would
remove at least some of the observed dependence of performance on sampling rate.
Nevertheless, this implementation performs quite well as it currently exists, as
we demonstrate in Chapter~\ref{chapter:Evaluation_and_Comparisons}.

\subsection{Evaluation Interface}
\label{subsection:Implementation-Evaluation_Interface}
The evaluation interface provides the means of evaluating a signal function
with inputs and producing effects in response to the signal function's outputs.
We would like to produce a set of constructs that interacts well with Haskell's
system for external IO.

The evaluation interface translates between signal functions and the standard
Haskell construct for sequencing effects and external inputs, namely,
{\em monads}~\cite{PeytonJones1993}. The inspiration for monads is drawn from
the rather esoteric domain of category theory, but the concept as applied to
programming languages is actually rather simple.

In a programming language with type constructors, a monad is simply a 1-arity
type constructor together with two operations. We shall call this type
constructor $m$. The first operation takes a value of any type
(call the type $a$) and produces a value of type $m \: a$ ($m$ applied to $a$).
The second operation takes a value of type $m \: a$ and a function from $a$ to
$m \: b$, and produces a $m \: b$. This means that this second operation,
monadic application, applies the function to the value in the context of the
monad.

For our purposes, there are two interesting properties of monads. One is that
for a specific type constructor, opaque primitives may be defined which
can be used, along with return, to build functions for the second argument
of the monadic application operation. In our case, we would like operations
to push an event to a signal function, to update the signal components of
the signal function, and to increment the time and sample the signal function.

This property is what enables monads to be used for input and output. A Haskell
{\tt main} program is an IO action. An IO action is a value of the monadic type
IO. This action is comprised of a primitive action (either a {\tt return} or a 
true IO primitive), possibly together with a function from the output of that
action to another IO action. Note that values with the IO type may appear
anywhere in a Haskell program, but they are only execute when they are returned
as a part of the {\tt main} IO program.

The Haskell typeclass for monads is defined:

\begin{code}
class Monad m where
  return :: a -> m a
  -- | Infix operator for monadic application
  (>>=)  :: m a -> (a -> m b) -> m b 
\end{code}

The other interesting property of monads is the existence of a class of monads
called monad transformers. These are arity-2 type constructors which take as
their first parameter a monad type constructor. The partial application of
such a type constructor to any monad type constructor then results in a monad.
A construct is then provided to lift values of the underlying monad to values
in the transforming monad, which may then be sequenced as normal. This leads
to the powerful concept of monad transformer stacks, where monads with many
different capabilities are combined without any more plumbing than explicit
{\tt lift} operations.

This property will be useful because we want our evaluation interface to be
part of a monad stack, the base of which is Haskell's IO monad. By formulating
the evaluation interface as a monad transformer, we need only define the
operations relevant to the evaluation of signal functions, and we can depend
on the constructs of the IO monad to interact with whatever inputs and outputs
are necessary. In some cases, we may not wish to use the IO monad at all (e.g
for testing or simulation). In this case, we can parameterize over another
monad, such as the Identity monad (which has no special operations and whose
context is just the value), or the State monad (which maintains an implicit 
state accessible by {\tt get} and {\tt put} operations).

The monad tranformer typeclass is defined as:

\begin{code}
class MonadTrans t where
  lift :: Monad m => m a -> t m a
\end{code}

Haskell also supports {\em do}-notation for monads. Do notation simply
makes monadic application implicit between lines and reverses the syntatic
order of lambda bindings, providing a way to use monads in a manner similar
to imperative programming. (See Figure~\ref{figure:monad-example} for an
example.)

\begin{figure}
\begin{code}
-- Do notation:
do putStr "Hello: "
   string <- getStr
   putStrLn $ "Nice to meet you " ++ string
   time <- getCurrentTime
   print time
   putStrLn "Look at the time! Bye."

-- Translates to:
putStr "Hello: " >> getStr >>= 
(\string -> (putStrLn $ "Nice to meet you " ++ string) >>
 getCurrentTime >>=
 (\time -> print time >> putStrLn "Look at the time! Bye."))
\end{code}
\hrule
\caption{An example of monadic $do$-notation and its equivalent monadic
expression.}
\label{figure:monad-example}
\end{figure}

The evaluation interface is exported as shown in
Section~\ref{section:System_Design_and_Interface-Evaluator}.

The {\tt SFEvalState} type constructor parameterizes over input types for signal
functions, and underlying monads for a monad transformer, but is not itself
a monad transformer. It describes the state of signal function evaluation.
It consists of a record with four members: the current signal function,
the set of handlers for outputs, the current input signal delta, and the last
sample time.

The {\tt SFEvalT} monad transformer is a newtype wrapper around the {\tt StateT}
monad available in the Haskell {\tt transformers} package. The StateT monad
wraps any other monad, providing a state which may be read using the {\tt get}
action and replaced with the {\tt put} action. An action in the underlying
monad is obtained by supplying a {\tt StateT} value and an initial state
value to the {\tt runStateT} function.

The {\tt SFEvalT} monad transformer does not make the {\tt get} and {\tt put}
actions available, but uses them in its implementation of the {\tt push},
{\tt update}, and {\tt sample} actions.

The {\tt push} action pushes an event onto the input of the signal function,
resulting in immediate evaluation of the relevant components signal function
and possible actuation of handlers specified for output events. It is
implemented by fetching the {\tt SFEvalState}, applying the event continuation
of the signal function contained by the {\tt SFEvalState} to the pushed event,
thus obtaining a new signal function and a list of events, applying the handlers
contained in the {\tt SFEvalState} to the output events, replacing the signal
function in the {\tt SFEvalState}, and replacing the {\tt SFEvalState}.

The {\tt update} action updates the current input signal delta, which will be
sampled at the next {\tt step} action. It has no immediately observable effect.
It simply updates the value of one signal in the input signal vector, a delta
of which is stored in the {\tt SFEvalState} record.

The {\tt step} action takes a time delta, and calls the signal continuation of
the stored signal function. It actuates the outputs using the stored handlers,
replaces the stored delta with an empty delta, and stores the resulting new
signal function in the state.

\section{Discussion}
\label{section:Discussion}

The system presented here, TimeFlies, demonstrates how using signal vectors
to type inputs and outputs enables push-based evaluation of events in a
signal-function system. We take advantage of this representation in several ways.

First, by separating components of inputs and outputs in the types, we are free
to create distinct, and often partial, representations of the input or output
of a signal function. This enables us to represent only the event occurrence
being pushed at that time.

Second, this separation also permits us to separate the process of gathering
the input to a signal function, and the process of handling its output, into
different points in a program. Using the evaluation interface described, an
event occurrence may be pushed onto one input of a signal function from one
point in a program (e.g. a mouse click handler), an input signal
may be updated in another (e.g. a mouse movement handler), and finally the
system may be sampled in a third place (e.g. an animation or audio timed
callback).

Finally, this approach enables further work on the implementation of the signal
function system to be separated from changes in the interface. By enabling
differing representations of the inputs and outputs of signal functions, we are
free to change these representations without the need to further constrain the
input and output types.

\section{Ongoing and Further Work}
\label{section:Ongoing_and_Further_Work}

TimeFlies, the system described here, has been implemented, but not
extensively tested. The immediate goal is to create a real-time application
which will permit a performance and implementation comparison of TimeFlies with
Yampa, the current state-of-the-art pull-based signal-function system.

In the future, we hope to apply run-time optimizations, using the technique used
for Yampa, to create a push-pull self-optimizing signal-function system. Further,
we hope to use this system as a basis for exploring signal-function FRP as a
basis for general-purpose application frameworks.

\section{Related Work}
\label{section:Related_Work}

Signal Function FRP was introduced as a model for Graphical User Interfaces~\cite{Courtney2001-1}.
The system was originally termed ``AFRP'' (Arrowized FRP). Yampa is a rewrite of
AFRP where signal functions apply a number of ad-hoc optimizatons to themselves
as they evolve. Yampa demonstrated a modest performance improvement
over AFRP~\cite{Nilsson2005}.

Reactive is a classic FRP system which implements push-based evaluation for events
by transforming behaviors to ``reactive~normal~form,'' where a behavior
is a non-reactive behavior running inside a switch, whose event stream carries
behaviors in reactive normal form. The system is evaluated by forking a Haskell
thread to repeatedly sample the non-reactive behavior, and then blocking on the
evaluation of the first occurrence in the event stream. When this occurrence
is yielded, the evaluation thread for the behavior is killed and a new
thread forked to evaluate the new behavior~\cite{Elliott2009}.

An alternate push-based system, based on {\em signal segments}, was presented in
work on the Curry-Howard correspondence between a form of temporal logic (LTL)
and FRP~\cite{Jeffrey2012}. However, this system continues to represent events
as option-valued signals. It therefore cannot be push-based in its response to
individual events, and cannot be push-pull, evaluating signals by pull-based
evaluation and events by push-based evaluation. It is push-based not in the
sense of evaluating events only when they occur, but rather in that it provides
a means to push signal segments (samples of a signal over intervals of time) to
the system and wait on its response.

\section{Conclusion}
\label{section:Conclusion}

We have presented TimeFlies, a system for push-pull signal-function Functional
Reactive Programming, and have shown how the use of a signal vectors as input
and output types for signal functions, together with GADT-based representations
of the inputs and outputs, permits the implementation of a push-pull system.

We have also described a general and flexible monadic evaluation interface for
TimeFlies, which permits us to interface the TimeFlies system with different
styles of IO systems, including multiple IO systems in the same application.

This opens up the exciting possibility that a signal-function FRP could become
an efficient and general framework for writing interactive applications.

\bibliographystyle{splncs}
\bibliography{thesis}

\end{document}