### 2012-11-27 Docs/IFL2012/conclusion.tex

\section{Conclusion}
\label{section:Conclusion}

We have presented TimeFlies, a system for push-pull signal-function Functional
Reactive Programming, and have shown how the use of a signal vectors as input
and output types for signal functions, together with GADT-based representations
of the inputs and outputs, permits the implementation of a push-pull system.

We have also described a general and flexible monadic evaluation interface for
TimeFlies, which permits us to interface the TimeFlies system with different
styles of IO systems, including multiple IO systems in the same application.

This opens up the exciting possibility that a signal-function FRP could become
an efficient and general framework for writing interactive applications.