### 2012-11-27 Docs/IFL2012/paper.tex

\documentclass[draft]{llncs}

\usepackage{fancyvrb}                                       % For code environment
\DefineVerbatimEnvironment{code}{Verbatim}{fontsize=\small} %

\title{Push-Pull Signal-Function Functional Reactive Programming}
\titlerunning{Push-Pull Signal-Function FRP}

\author{Edward Amsden}
\institute{Rochester Institute of Technology\\\email{eca7215@cs.rit.edu}}

\begin{document}
\maketitle

\input{abstract.tex}
\input{introduction.tex}
\input{design.tex}
\input{implementation.tex}
\input{discussion.tex}
\input{related.tex}
\input{conclusion.tex}

\bibliographystyle{splncs}
\bibliography{thesis}

\end{document}