### 2012-11-29 Docs/IFL2012/abstract.tex
edits to ifl abstract and intro

\begin{abstract}
Functional Reactive Programming is a promising model for writing interactive and
time-dependent programs. Signal-function FRP is a variant of FRP which provides
advantages in modularity and correctness, but has proven difficult to implement
efficiently.

One important implementation trade-off is whether evaluation of an FRP program
is push-based or pull-based. Signal vectors are an abstraction which addresses
this by providing a type-level description of the inputs and outputs of signal
functions, distinguishing components that benefit from a push-based
implementation from those that benefit from a pull-based implementation. This
makes it possible to combine both implementation strategies in a single system.

We describe a signal-function FRP system which provides push-based evaluation
for events, pull-based evaluation for signals, and a simple monadic evaluation
interface that permits the system to easily integrate with one or more IO
systems.
\end{abstract}