### 2013-01-29 Docs/Thesis/introduction.tex

\chapter{Introduction}
\label{chapter:Introduction}

Most of the useful programs which programmers are asked to write must react to inputs which are not available at the time the program starts,
and produce effects at many different times throughout the execution of the program. Examples of these programs include GUI applications,
web browsers, web applications, video games, multimedia authoring and playback tools, operating system shells and kernels, servers,
robotic control programs, and many others. This attribute of a program is called reactivity.

Functional reactive programming (FRP) is a promising class of abstractions for
writing interactive programs in functional languages. The FRP abstractions are
{\em behaviors} (also termed {\em signals}), which are functions of continuous
time, and {\em events}, which are sequences of time-value pairs. These
abstractions translate desirable properties of the underlying functional
languages, such as higher-order functions and referential transparency, to
reactive constructs, generally without modifications to the underlying language.
This permits the compositional construction of reactive programs in purely
functional languages.

Functional reactive programming was introduced, though not quite by name, with
Fran~\cite{Elliott1997}, a compositional system for writingm reactive
animations. From the start, two key challenges emerged: the enforcement of
{\em causality}\footnote{Causality is the property that a value in an FRP
program depends only on present and past values. Obviously, a program which
violates this property is impossible to evaluate, but in some FRP systems such
programs are unfortunately quite easy to write down.} in FRP programs, and the
efficient evaluation of FRP programs.

The first challenge was addressed by representing FRP programs, not as
compositions of behaviors and events, but as compositions of transformers of
behaviors and events called {\em signal functions}. By not permitting direct
access to behaviors or events, but representing them implicitly, signal
functions prohibit accessing past or future time values, only allowing access to
values in the closure of the signal function and current input values. Signal
function FRP programs are written by composing signal functions, and are
evaluated using a function which provides input to the signal function and acts
on the output values. This evaluation function is generally elided in existing
literature on signal functions, but we provide a documented and improved
interface for evaluating signal functions. A further advantage of
signal-function FRP programs is that they are more readily composed,
since additional signal functions may be composed on the input or output side of
a signal function, rather than simply the output side as in classic
(non-signal-function) FRP systems.

The second challenge was addressed for classic FRP by the creation of a hybrid
push-pull FRP system~\cite{Elliott2009}. This system relied on the distinction
between reactivity and time dependence to decompose behaviors into phases of
constant or simply time-dependent values, with new phases initiated by event
occurrences. However, this implementation did not address concerns of causality
or composability. Further, the existence of events as first-class values in
classic FRP forced the use of an impure and resource-leaking technique to
compare events when merging, in order to determine which event had the first
occurrence after the current time. Further, since the classic FRP interface
permits access to only the output of a behavior or event, and is always bound to
its implicit inputs, the best the push-based implementation can do is to block
when no computation needs to be performed. The computation cannot be suspended
entirely as a value representation.

To address both challenges in a FRP implementation, we combine the approaches
and demonstrate a push-pull signal-function FRP system. The current
implementation approach for signal-function FRP is inherently
pull-based~\cite{Nilsson2002}, but recent work on N-ary
FRP~\cite{Sculthorpe2011}, a variant of signal-function FRP which enumerates
inputs and distinguishes between events and behaviors, provides a representation
which suggests a method of push-based evaluation. In previous implementations of
signal-function FRP, events were simply option-valued signals. This approach has
the advantage of being simple to implement. However, it does not have an obvious
semantics for event merging. Further, it necessitates pull-based evaluation,
since there is no way to separate the events which may be pushed from the truly
continuous values which must be repeatedly sampled.

This thesis describes the design and implementation of TimeFlies, a push-pull signal-function FRP system. This system is presented as a library
of combinators for constructing and evaluating FRP programs.