### 2013-07-17 Docs/Thesis/implementation.tex

\chapter{Implementation}
\label{chapter:Implementation}

Having established the design and semantics of the reactive system, I present a purely 
functional implementation. For this implementation, I will make significant use of
Generalized Algebraic Datatypes (GADTs)~\cite{Cheney2003,Xi2003}. GADTs provide a mechanism
for expressing abstractions which are not easily expressible in a standard
Hindley-Milner type system, while maintaining the type soundness of Haskell.
In particular, GADTs provide the power we require to take advantage of the
signal vector representation of signal function inputs and outputs.

\input{implementation-signalfunctions}

\input{implementation-evalinterface}